### 2022-09-22 Paper.tex
Initial Overleaf Import

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{Digital Authoritarianism and Western Foreign Aid}
\author{Nicholas Ray}
\date{September 2022}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\doublespacing
\usepackage[backend=biber, style=authoryear, maxbibnames=99]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{References.bib}
\begin{document}
\maketitle
\section*{What needs to be explained} 
China and U.S. in foreign aid competition, with large amounts of money directed toward internet related infrastructure.
Understand why China wants to give but why the U.S.? And why internet infrastructure? And why now?

\section*{Literature Review}

\section*{Theory} 
U.S. wants to protect American Big Tech markets that are potentially going to be lost in a zero-sum game with Chinese companies (if China successfully shares tech and builds infrastructure).

\textbf{Implications} 
\begin{align*}
    H_1a:\; & \text{A pattern should be evident of U.S. money flowing into places where either}\\
    & \text{Big Tech markets are OR}\\
    H_1b:\; & \text{Where Chinese influence/money is increasing OR}\\
    H_2:\; & \text{Where Chinese influence is NOT increasing (U.S. defending markets)}
\end{align*}

\section*{Alternative Explanations}
\textbf{Theory}

The U.S. is not concerned necessarily about Big Tech markets, but instead increasing aid in an attempt to thwart China's goals of economic growth abroad.

\textbf{Implication}
\begin{align*}
    H_3:\; & \text{There should be a relatively random pattern of spending or at least}\\
    & \text{a pattern not associated with Big Tech markets}
\end{align*}

\textbf{Theory}

The U.S. is not necessarily concerned about market access for American Big Tech or China's goal of increasing economic growth through trade/economic diplomacy, but rather trying to counter Chinese money's influence on digital authoritarianism. The U.S. would then be trying to share it's own tech and build it's own internet infrastructure to have more control over the recipient's internet freedom levels.

\textbf{Implication}
\begin{align*}
    H_4:\; & \text{U.S. spending patters would be associated with increasing}\\
    & \text{internet freedom levels and potentially concentrated in places that are}\\
    & \text{not already receiving tech and infrastructure from China}\\
\end{align*}

\section*{Methodology}
Test? Right now I'm thinking of taking data on U.S., Chinese money, internet/social media usage stats and regressing U.S. money on Chinese money and internet/social media usage to try and discern any pattern. Obvious OVB, Simultaneity issues, but we'll keep thinking about it. And note the fact that this is likely a time series (actually panel data), so OLS not satisfactory.

\section*{Complications}
Data on foreign aid is outdated or only on U.S. and China. (Not G7)

Theoretically, it's difficult to conceive of how to differentiate these patterns and hypotheses.

\section*{One-Pager}
The People's Republic of China has been an enduring act on the world stage of the last ten years. Unveiling the Belt and Road Initiative (BRI) in 2013 and then the Digital Silk Road (DSR) in 2015, China has become the world's leading financier, surpassing even the United States in yearly foreign aid expenditures (\cite{dreherND}).
Investing in the infrastructure of countries across the globe, it is widely believed that these projects are motivated by both the desire for influence and economic growth, with Chinese companies enjoying unprecedented access to international markets.

Remarkably, the U.S. has decided to act. Surrounded by other members of the G7 at their recent summer summit, President Biden announced that \$600 billion will be spent by the member countries over the next five years to explicitly counter China's intentions (\cite{NPR}). Most interestingly, at least a billion of those dollars are officially earmarked for building internet infrastructure. \nocite{US} \nocite{DSR}

It is not ex ante obvious why the U.S. has taken such a keen and expensive interest in competing with Chinese money or why the internet is a significant focus. China, on the other hand, clearly has much to gain from it's behavior given the domestic importance of economic growth and the lure of becoming a top contender for world power. But does the U.S. really feel a similar demand to grow it's political and financial empire? And even if that is the case, why now?

I argue that the U.S. is trying to combat a perceived increase in digital authoritarianism led by Chinese funding and technology while protecting American Big Tech's access to international markets. I speculate that the U.S. views the use of the internet in other countries as a zero-sum game, where the adoption of Chinese tools and infrastructure prevents U.S.-based companies from flourishing. A clue suggesting this new perspective was a declaration made just two months prior to the G7 spending announcement, where the U.S. and 55 other countries committed to push for a democratic use of the internet worldwide (\cite{Declaration}).

Should this explanation for planned G7 foreign aid be true, it implies that Western money should be spent in an observable pattern consistent with these goals. Specifically, current Chinese foreign aid and the location of profitable technology markets should be strong predictors of future Western foreign aid. The comparison of Western aid over time would thus be the goal of this research, with the expectation that aid increases in countries with strong consumer bases and high exposure to Chinese impact.

The substantive implication of this explanation, should it not be proven false, is the realization that technology and the internet have become central considerations for international politics. Foreign aid, to the extent that it is a strategic endeavor, could be partly determined by cybersecurity concerns and the ability for private firms to profit off of internet activity- motivations that have been largely unconsidered by scholars. A better understanding of these technological incentives might prove salubrious for the predictive power of political science as the internet and AI continue to grow in prominence.
\printbibliography
\end{document}

### 2022-09-22 Plan.tex
Plan

Goal is to create a theory that rationalizes the U.S./G7's decision to increase aid to internet-related infrastructure (i.e., the U.S. says it is a value-driven effort to increase democratic democracy, but how is it rational for the U.S. to invest in that?) Counter-point: why is it surprising that donor countries, allegedly interested in economic development, would make internet infrastructure part of that development goal since the internet is good for the economy? That's certainly possible and could be one of my hypotheses... can I get information on the type of development most needed in recipient countries and use that to make a case that the internet should not be a priority?

Need to find extant theories in the literature on how U.S./other countries rationally distribute aid and see where my theory can contribute or not

Then, would need some way to grapple with the empirical problem or supporting any theoretical implications

### 2022-09-22 Plan.tex
hypotheses

@@
 Then, would need some way to grapple with the empirical problem or supporting any theoretical implications
+
+We need to be more innocent about this project.
+We really don't know what is going on.
+There seem to be some alternative hypotheses: internet infrastructure good for recipient country and donors altruistic, recipients declare that is a need and donors responsive whether self-interested or not.
+Our hypothesis: donor countries interested in internet infrastructure to prevent China from blocking/subverting West's digital access in these countries...

### 2022-09-22 Plan.tex
changes

@@
 There seem to be some alternative hypotheses: internet infrastructure good for recipient country and donors altruistic, recipients declare that is a need and donors responsive whether self-interested or not.
-Our hypothesis: donor countries interested in internet infrastructure to prevent China from blocking/subverting West's digital access in these countries...
+Our hypothesis: donor countries interested in internet infrastructure to prevent China from overtaking West's digital access in these countries...

### 2022-09-22 Paper.tex
Zotero
LitReview

@@
 
-\section*{Literature Review}
+\section*{Determinants of Aid}
+\subsection*{Self-Interest versus Altruism}
+Hoeffler, Outram: ``Our results indicate that all bilateral donors allocate aid according to their self-interest and recipient need''
+
+Younas: Trade seems to account for donor behavior, along with priorities in human rights and miseries rather than poverty.
+
+Wiseman, Young: Donors base behavior off of ``warm-glow''
 

### 2022-09-28 OldTheory.tex
Clean

--- 2022-09-22 Paper.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\title{Digital Authoritarianism and Western Foreign Aid}
-\author{Nicholas Ray}
-\date{September 2022}
-\usepackage[margin=1in]{geometry}
-\usepackage{mathtools,amssymb,amsthm}
-\usepackage{setspace}
-\doublespacing
-\usepackage[backend=biber, style=authoryear, maxbibnames=99]{biblatex}
-\renewbibmacro{in:}{}
-\renewbibmacro*{volume+number+eid}{%
-  \printfield{volume}
-  \setunit*{\addnbthinspace}
-  \printfield{number}
-  \setunit{\addcomma\space}
-  \printfield{eid}}
-\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
-\AtEveryBibitem{
-  \clearfield{issn}
-  \clearfield{month}
-  \clearfield{urlyear}
-  \clearlist{language}
-  \clearfield{note}
-  \ifentrytype{online}{}{
-    \clearfield{url}
-  }
-}
-\addbibresource{References.bib}
-\begin{document}
-\maketitle
-\section*{What needs to be explained} 
-China and U.S. in foreign aid competition, with large amounts of money directed toward internet related infrastructure.
-Understand why China wants to give but why the U.S.? And why internet infrastructure? And why now?
-
-\section*{Literature Review}
-
 \section*{Theory} 
@@
 
-\textbf{Implications} 
 \begin{align*}
@@
 \section*{Alternative Explanations}
-\textbf{Theory}
-
 The U.S. is not concerned necessarily about Big Tech markets, but instead increasing aid in an attempt to thwart China's goals of economic growth abroad.
@@
 \end{align*}
-
-\textbf{Theory}
 
@@
 \end{align*}
-
-\section*{Methodology}
-Test?
-Right now I'm thinking of taking data on U.S., Chinese money, internet/social media usage stats and regressing U.S.
-money on Chinese money and internet/social media usage to try and discern any pattern.
-Obvious OVB, Simultaneity issues, but we'll keep thinking about it.
-And note the fact that this is likely a time series (actually panel data), so OLS not satisfactory.
-
-\section*{Complications}
-Data on foreign aid is outdated or only on U.S. and China. (Not G7)
-
-Theoretically, it's difficult to conceive of how to differentiate these patterns and hypotheses.
-
-\section*{One-Pager}
-The People's Republic of China has been an enduring act on the world stage of the last ten years.
-Unveiling the Belt and Road Initiative (BRI) in 2013 and then the Digital Silk Road (DSR) in 2015, China has become the world's leading financier, surpassing even the United States in yearly foreign aid expenditures (\cite{dreherND}).
-Investing in the infrastructure of countries across the globe, it is widely believed that these projects are motivated by both the desire for influence and economic growth, with Chinese companies enjoying unprecedented access to international markets.
-
-Remarkably, the U.S.
-has decided to act.
-Surrounded by other members of the G7 at their recent summer summit, President Biden announced that \$600 billion will be spent by the member countries over the next five years to explicitly counter China's intentions (\cite{NPR}).
-Most interestingly, at least a billion of those dollars are officially earmarked for building internet infrastructure.
-\nocite{US} \nocite{DSR}
-
-It is not ex ante obvious why the U.S.
-has taken such a keen and expensive interest in competing with Chinese money or why the internet is a significant focus.
-China, on the other hand, clearly has much to gain from it's behavior given the domestic importance of economic growth and the lure of becoming a top contender for world power.
-But does the U.S.
-really feel a similar demand to grow it's political and financial empire?
-And even if that is the case, why now?
-
-I argue that the U.S.
-is trying to combat a perceived increase in digital authoritarianism led by Chinese funding and technology while protecting American Big Tech's access to international markets.
-I speculate that the U.S.
-views the use of the internet in other countries as a zero-sum game, where the adoption of Chinese tools and infrastructure prevents U.S.-based companies from flourishing.
-A clue suggesting this new perspective was a declaration made just two months prior to the G7 spending announcement, where the U.S.
-and 55 other countries committed to push for a democratic use of the internet worldwide (\cite{Declaration}).
-
-Should this explanation for planned G7 foreign aid be true, it implies that Western money should be spent in an observable pattern consistent with these goals.
-Specifically, current Chinese foreign aid and the location of profitable technology markets should be strong predictors of future Western foreign aid.
-The comparison of Western aid over time would thus be the goal of this research, with the expectation that aid increases in countries with strong consumer bases and high exposure to Chinese impact.
-
-The substantive implication of this explanation, should it not be proven false, is the realization that technology and the internet have become central considerations for international politics.
-Foreign aid, to the extent that it is a strategic endeavor, could be partly determined by cybersecurity concerns and the ability for private firms to profit off of internet activity- motivations that have been largely unconsidered by scholars.
-A better understanding of these technological incentives might prove salubrious for the predictive power of political science as the internet and AI continue to grow in prominence.
-\printbibliography
-\end{document}

### 2022-09-28 Paper.tex
Clean

--- 2022-09-22 Paper.tex
@@
 \maketitle
-\section*{What needs to be explained} 
-China and U.S. in foreign aid competition, with large amounts of money directed toward internet related infrastructure.
-Understand why China wants to give but why the U.S.? And why internet infrastructure? And why now?
+\section*{Introduction}
+(What needs to be explained? What is the puzzling observation? What is the research question?)
+
+Evidence that internet freedom has been a part of U.S. foreign policy for the past decade, but no empirical evidence that U.S. foreign aid 
+(an important part of U.S.
+foreign policy) is driven by internet freedom concerns.
+If it is, why is it rational for the U.S.
+to be concerned about internet freedom?
+(Further, maybe, does U.S.
+aid have a positive effect on internet freedom?) I argue that to the extent that U.S.
+aid is related to internet freedom, it is rational for the U.S.
+to expend resources on increasing internet freedom because 1) it secures them access for the future to developing country's markets that could give profitable personal data to U.S.
+Big Tech and 2) the U.S.
+has an interest in thwarting China's efforts at increasing domestic economic growth through their technology companies.
 
 \section*{Literature Review}
+\subsection*{Determinants of Aid}
+\textbf{Self-Interest versus Altruism}
+
+Hoeffler, Outram: ``Our results indicate that all bilateral donors allocate aid according to their self-interest and recipient need''
+
+Younas: Trade seems to account for donor behavior, along with priorities in human rights and miseries rather than poverty.
+
+Wiseman, Young: Donors base behavior off of ``warm-glow''
+
+National interests versus recipient need (general debate about self-interest or in the interest of recipients).
 
 \section*{Theory}
-U.S. wants to protect American Big Tech markets that are potentially going to be lost in a zero-sum game with Chinese companies (if China successfully shares tech and builds infrastructure).
-
-\textbf{Implications} 
-\begin{align*}
-    H_1a:\; & \text{A pattern should be evident of U.S. money flowing into places where either}\\
-    & \text{Big Tech markets are OR}\\
-    H_1b:\; & \text{Where Chinese influence/money is increasing OR}\\
-    H_2:\; & \text{Where Chinese influence is NOT increasing (U.S. defending markets)}
-\end{align*}
-
-\section*{Alternative Explanations}
-\textbf{Theory}
-
-The U.S. is not concerned necessarily about Big Tech markets, but instead increasing aid in an attempt to thwart China's goals of economic growth abroad.
-
-\textbf{Implication}
-\begin{align*}
-    H_3:\; & \text{There should be a relatively random pattern of spending or at least}\\
-    & \text{a pattern not associated with Big Tech markets}
-\end{align*}
-
-\textbf{Theory}
-
-The U.S.
-is not necessarily concerned about market access for American Big Tech or China's goal of increasing economic growth through trade/economic diplomacy, but rather trying to counter Chinese money's influence on digital authoritarianism.
-The U.S.
-would then be trying to share it's own tech and build it's own internet infrastructure to have more control over the recipient's internet freedom levels.
-
-\textbf{Implication}
-\begin{align*}
-    H_4:\; & \text{U.S. spending patters would be associated with increasing}\\
-    & \text{internet freedom levels and potentially concentrated in places that are}\\
-    & \text{not already receiving tech and infrastructure from China}\\
-\end{align*}
 
 \section*{Methodology}
-Test?
-Right now I'm thinking of taking data on U.S., Chinese money, internet/social media usage stats and regressing U.S.
-money on Chinese money and internet/social media usage to try and discern any pattern.
-Obvious OVB, Simultaneity issues, but we'll keep thinking about it.
-And note the fact that this is likely a time series (actually panel data), so OLS not satisfactory.
 
 \section*{Complications}
-Data on foreign aid is outdated or only on U.S. and China. (Not G7)
 
-Theoretically, it's difficult to conceive of how to differentiate these patterns and hypotheses.
+\section*{Conclusion}
 
-\section*{One-Pager}
-The People's Republic of China has been an enduring act on the world stage of the last ten years.
-Unveiling the Belt and Road Initiative (BRI) in 2013 and then the Digital Silk Road (DSR) in 2015, China has become the world's leading financier, surpassing even the United States in yearly foreign aid expenditures (\cite{dreherND}).
-Investing in the infrastructure of countries across the globe, it is widely believed that these projects are motivated by both the desire for influence and economic growth, with Chinese companies enjoying unprecedented access to international markets.
-
-Remarkably, the U.S.
-has decided to act.
-Surrounded by other members of the G7 at their recent summer summit, President Biden announced that \$600 billion will be spent by the member countries over the next five years to explicitly counter China's intentions (\cite{NPR}).
-Most interestingly, at least a billion of those dollars are officially earmarked for building internet infrastructure.
-\nocite{US} \nocite{DSR}
-
-It is not ex ante obvious why the U.S.
-has taken such a keen and expensive interest in competing with Chinese money or why the internet is a significant focus.
-China, on the other hand, clearly has much to gain from it's behavior given the domestic importance of economic growth and the lure of becoming a top contender for world power.
-But does the U.S.
-really feel a similar demand to grow it's political and financial empire?
-And even if that is the case, why now?
-
-I argue that the U.S.
-is trying to combat a perceived increase in digital authoritarianism led by Chinese funding and technology while protecting American Big Tech's access to international markets.
-I speculate that the U.S.
-views the use of the internet in other countries as a zero-sum game, where the adoption of Chinese tools and infrastructure prevents U.S.-based companies from flourishing.
-A clue suggesting this new perspective was a declaration made just two months prior to the G7 spending announcement, where the U.S.
-and 55 other countries committed to push for a democratic use of the internet worldwide (\cite{Declaration}).
-
-Should this explanation for planned G7 foreign aid be true, it implies that Western money should be spent in an observable pattern consistent with these goals.
-Specifically, current Chinese foreign aid and the location of profitable technology markets should be strong predictors of future Western foreign aid.
-The comparison of Western aid over time would thus be the goal of this research, with the expectation that aid increases in countries with strong consumer bases and high exposure to Chinese impact.
-
-The substantive implication of this explanation, should it not be proven false, is the realization that technology and the internet have become central considerations for international politics.
-Foreign aid, to the extent that it is a strategic endeavor, could be partly determined by cybersecurity concerns and the ability for private firms to profit off of internet activity- motivations that have been largely unconsidered by scholars.
-A better understanding of these technological incentives might prove salubrious for the predictive power of political science as the internet and AI continue to grow in prominence.
+\pagebreak
 \printbibliography

### 2022-09-28 Plan.tex
Clean

@@
 can I get information on the type of development most needed in recipient countries and use that to make a case that the internet should not be a priority?
+
+What about USaid~Chinese aid + internet freedom levels?
+If USaid theoretically concerned with digital autocracy.
+See if there's a precedent for USaid being concerned with digital democracy.
+Use most recent data I can (e.g., OECD 2021) to see if during the Biden administration there has already been evidence of this despite majority of funds not being actualized yet (announcement in 2022).
 
@@
 Our hypothesis: donor countries interested in internet infrastructure to prevent China from overtaking West's digital access in these countries...
+
+Debate in literature on if aid is self-interested, altruistic... If US was aiding digital democracy, is that a self-interested or altruistic act?

### 2022-09-28 Paper.tex
progress

@@
 (What needs to be explained? What is the puzzling observation? What is the research question?)
+Has China's aid project changed how the United States conducts it's aid?
+I argue that we should expect the Chinese project to have an effect because the United States has made internet freedom part of it's foreign aid policy, which is perhaps being affected by Chinese aid itself.
+Been a lot of interest recently on the effects of China's BRI (e.g., cite some highly published journals).
 
-Evidence that internet freedom has been a part of U.S. foreign policy for the past decade, but no empirical evidence that U.S. foreign aid 
-(an important part of U.S.
-foreign policy) is driven by internet freedom concerns.
-If it is, why is it rational for the U.S.
-to be concerned about internet freedom?
-(Further, maybe, does U.S.
-aid have a positive effect on internet freedom?) I argue that to the extent that U.S.
+Evidence that internet freedom has been a part of U.S.
+foreign policy for the past decade, but no observational evidence that U.S.
+foreign aid (an important part of U.S.
+foreign policy) is related to the internet freedom of recipient countries.
+Further, it is not ex ante obvious why the U.S.
+should rationally be spending money on a relatively abstract concept of internet freedom (literature on how concern for democracy drives U.S.?
+There's a lot of literature on if the US's aid program is self-interested or altruistic, which can leverage here (i.e., given that U.S.
+seems to be self-interested, why?))
+
+I argue that to the extent that U.S.
 aid is related to internet freedom, it is rational for the U.S.
 to expend resources on increasing internet freedom because 1) it secures them access for the future to developing country's markets that could give profitable personal data to U.S.
-Big Tech and 2) the U.S.
+Big Tech (there is a possibility that internet is a zero-sum game between China, U.S., where U.S.
+wants to preempt China's takeover of technology markets, i.e.
+technological arms race, access war) and 2) the U.S.
 has an interest in thwarting China's efforts at increasing domestic economic growth through their technology companies.
@@
 \section*{Theory}
+It is rational for the U.S. to (condition?) spend foreign aid according to internet freedom levels of recipient because...
 

### 2022-09-29 Paper.tex
bib

--- 2022-09-28 Paper.tex
@@
 \usepackage[utf8]{inputenc}
-\title{Digital Authoritarianism and Western Foreign Aid}
+\title{\vspace{-2cm}Digital Authoritarianism and United States Foreign Aid\vspace{-0.5cm}}
 \author{Nicholas Ray}
-\date{September 2022}
+\date{\vspace{-0.5cm}September 2 2022\vspace{-1cm}}
 \usepackage[margin=1in]{geometry}
@@
 }
-\addbibresource{References.bib}
+\addbibresource{ForeignAidBib.bib}
 \begin{document}
@@
 \section*{Introduction}
-(What needs to be explained? What is the puzzling observation? What is the research question?)
-
-Evidence that internet freedom has been a part of U.S. foreign policy for the past decade, but no empirical evidence that U.S. foreign aid 
-(an important part of U.S.
-foreign policy) is driven by internet freedom concerns.
-If it is, why is it rational for the U.S.
-to be concerned about internet freedom?
-(Further, maybe, does U.S.
-aid have a positive effect on internet freedom?) I argue that to the extent that U.S.
-aid is related to internet freedom, it is rational for the U.S.
-to expend resources on increasing internet freedom because 1) it secures them access for the future to developing country's markets that could give profitable personal data to U.S.
-Big Tech and 2) the U.S.
-has an interest in thwarting China's efforts at increasing domestic economic growth through their technology companies.
+U.S.'s history of concern over global internet freedom. (Brookings, Rand, Congressional Bill, Declaration, Partnership for Global Infrastructure and Investment).
+Given the U.S.'s history of concern over global internet freedom, how has U.S. foreign policy changed in response to China's involvement in developing countries? 
+While there has been a wealth of academic attention on the potential effects of China's BRI (e.g., good journal articles), few have focused on how it might have shifted U.S.
+foreign policy.
+It is reasonable to expect that China's program would have \textit{some} effect on U.S.
+policy since 1) China's program is drawing attention from U.S.
+(Partnership, Declaration; and it could potentially be negatively affecting internet freedom) and 2) research has shown that U.S.
+aid is effective in promoting democracy despite Chinese aid prevalence (Blair, Marty, Roessler 2021).
 
 \section*{Literature Review}
-\subsection*{Determinants of Aid}
-\textbf{Self-Interest versus Altruism}
+Most foreign aid allocated to self-interest and not recipient need (\cite{hoeffler_need_2011})
 
-Hoeffler, Outram: ``Our results indicate that all bilateral donors allocate aid according to their self-interest and recipient need''
+Aid ineffectiveness in achieving economic growth or promoting democratic institutions (Chong, Gradstein, Calderon)
 
-Younas: Trade seems to account for donor behavior, along with priorities in human rights and miseries rather than poverty.
+Aid from U.S. successful in promoting democracy (Blair, Marty, Roessler 2021)
 
-Wiseman, Young: Donors base behavior off of ``warm-glow''
+Aid from U.S.
+tends to be more effective in countries that receive no substantial support from China (so we should see funding directed towards countries lacking Chinese aid already)(Dreher et al 2021 AER)
 
-National interests versus recipient need (general debate about self-interest or in the interest of recipients).
+Self-interest (Alesina and Dollar)
 
-\section*{Theory}
+Chinese aid associated with increase in authoritarian values; World bank aid associated with democracy (Gehring, Kaplan, and Wong)
+
+aid increases democracy (Dietrich, Wright)
+
+Commercial self-interest present but possibly overblown (Dreher, Nunnenkamp, Thiele)
+
+Bilateral aid more likely to be conditioned on things and less impartial, implying that U.S. should use it more if it is really concerned with internet freedom changes (Christensen, Homer, Nielson)
+
+Trade benefits drive aid (Younas)
+
+Political economic factors can shape preferences for foreign aid in U.S. (Milner, Tingley)
 
 \section*{Methodology}
+Dependent variable: US Aid
+Main explanatory variables: Chinese Aid, Internet Freedom
+Controls (according to Cinelli et al.): GDP per capita (recipient need), Internet Usership (Internet development indicators), Political Freedom (possible confounder)
 
+Chinese data set (Custer et al. 2021; Dreher et al. 2022 (Banking on Beijing))
+
+If concerned about simultaneity between US aid and Internet Freedom, maybe use instrument related to Internet Freedom but not to US aid. 
 \section*{Complications}
@@
 \section*{Conclusion}
+(Speculate about why US might be rationally interested in increasing internet freedom) Given the literature on the self-interestedness of donors, if U.S.
+aid is changing in response to Chinese aid, what might be a rational motivation?
+Basically: aid effectually increases democracy, democracy is good for business (American Big Tech).
 

### 2022-09-29 TheoryIdeas.tex
Updated

--- 2022-09-28 OldTheory.tex
@@
 \end{align*}
+
+9/28
+If U.S.
+foreign aid policy is rational and related to internet freedom, then the U.S.
+must stand to gain from spending money on improving others' internet freedom.
+I argue that the U.S.
+benefits in two main ways from supporting internet freedom: protecting American technology's access to developing markets (possibly being encouraged by China to use Chinese technology instead) while ensuring that China faces resistance in continuing its economic growth (which is it's main goal of this project).
+
+If this is true, then we should observe U.S. foreign aid spending patterns with one set of the following hypotheses:
+
+\begin{align*}
+    H_1a:\; & \text{U.S. aid should be associated with countries with high levels of internet freedom.}\\
+    H_1b:\; & \text{U.S. aid should be associated with countries with low levels of internet freedom.}\\
+    H_2a:\; & \text{U.S. aid should be associated with countries receiving high amounts of Chinese aid.}\\
+    H_2b:\; & \text{U.S. aid should be associated with countries receiving low amounts of Chinese aid.}
+\end{align*}
+
+9/29
+\section*{Theory}
+Depending on how the literature review goes, we can mention our theory of why it is rational for the U.S.
+to engage in promoting internet freedom (aid effectually increases democracy, and democracy is good for business (American Big Tech)).

### 2022-09-30 Paper.tex
Progress

@@
 \doublespacing
-\usepackage[backend=biber, style=authoryear, maxbibnames=99]{biblatex}
+\usepackage[backend=biber, style=authoryear, maxbibnames=99,uniquelist=false]{biblatex}
 \renewbibmacro{in:}{}
@@
   \clearfield{note}
+  \clearfield{month}
+  \clearfield{day}
   \ifentrytype{online}{}{
@@
 \section*{Introduction}
-U.S.'s history of concern over global internet freedom. (Brookings, Rand, Congressional Bill, Declaration, Partnership for Global Infrastructure and Investment).
-Given the U.S.'s history of concern over global internet freedom, how has U.S. foreign policy changed in response to China's involvement in developing countries? 
-While there has been a wealth of academic attention on the potential effects of China's BRI (e.g., good journal articles), few have focused on how it might have shifted U.S.
+The United States (U.S.) has been openly concerned about the internet since at least 1997, when President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (\cite{government1997}).
+By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as internet freedom programs cropped up in different facets of the U.S.
+government, such as the Department of State (e.g., \cite{government2021}) and the U.S.
+Agency for Global Media (\cite{government2022d}).
+U.S.
+attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).
+
+Today internet freedom has seemingly reached a crescendo in U.S.
 foreign policy.
-It is reasonable to expect that China's program would have \textit{some} effect on U.S.
-policy since 1) China's program is drawing attention from U.S.
-(Partnership, Declaration; and it could potentially be negatively affecting internet freedom) and 2) research has shown that U.S.
-aid is effective in promoting democracy despite Chinese aid prevalence (Blair, Marty, Roessler 2021).
+Bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S.
+and over 60 other countries to combat ``digital authoritarianism'' (\cite{government2022b}), and an unprecedented, U.S.-led G7 spending plan was unveiled to improve physical and digital infrastructure abroad (\cite{government2022c})- all in 2022.
+
+Along the way, the People's Republic of China (China) and the Russian Federation (Russia) have been named the antagonists to global internet freedom and the explicit motivations behind the U.S.'s focus on the subject.
+During this intensifying period of U.S.
+apprehension of digital authoritarianism, China launched it's Belt and Road Initiative (BRI) in 2013 and the accompanying Digital Silk Road Initiative (DSR) in 2015.
+Collectively, these projects have catalyzed over one trillion U.S.
+dollars in developing-country investment with the intent to improve China's perception and influence in the world while fostering it's domestic economic growth (\cite{dreher2022}).
+
+Ample academic work has recently attempted to outline the effects of China's aid program (e.g., \cite{gehring2022}; \cite{dreher2018}; \cite{blair2021}).
+Yet, little research has explored the potential effects of China's spending on U.S.
+foreign policy.
+In this paper, I seek to understand if U.S.
+foreign aid, an important aspect of foreign policy, can be understood as a response to China's aid ambitions.
+Specifically, I hypothesize that U.S.
+foreign aid to Africa is related to a recipient's internet freedom level and their influx of Chinese aid.
 
 \section*{Literature Review}
-Most foreign aid allocated to self-interest and not recipient need (\cite{hoeffler_need_2011})
+Most foreign aid allocated to self-interest and not recipient need (\cite{hoeffler2011})
 
@@
 
-Aid from U.S. successful in promoting democracy (Blair, Marty, Roessler 2021)
+Aid from U.S. successful in promoting democracy (\cite{blair2022})
 
@@
 Dependent variable: US Aid
+
 Main explanatory variables: Chinese Aid, Internet Freedom
+
 Controls (according to Cinelli et al.): GDP per capita (recipient need), Internet Usership (Internet development indicators), Political Freedom (possible confounder)
@@
 If concerned about simultaneity between US aid and Internet Freedom, maybe use instrument related to Internet Freedom but not to US aid.
+
+DiD? With two countries that have similar levels of Internet Freedom and Chinese aid (parallel trends) and then both experience increases in Chinese aid?
 \section*{Complications}

### 2022-10-01 Extra.tex
Progress

--- 2022-09-28 Paper.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\title{Digital Authoritarianism and Western Foreign Aid}
-\author{Nicholas Ray}
-\date{September 2022}
-\usepackage[margin=1in]{geometry}
-\usepackage{mathtools,amssymb,amsthm}
-\usepackage{setspace}
-\doublespacing
-\usepackage[backend=biber, style=authoryear, maxbibnames=99]{biblatex}
-\renewbibmacro{in:}{}
-\renewbibmacro*{volume+number+eid}{%
-  \printfield{volume}
-  \setunit*{\addnbthinspace}
-  \printfield{number}
-  \setunit{\addcomma\space}
-  \printfield{eid}}
-\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
-\AtEveryBibitem{
-  \clearfield{issn}
-  \clearfield{month}
-  \clearfield{urlyear}
-  \clearlist{language}
-  \clearfield{note}
-  \ifentrytype{online}{}{
-    \clearfield{url}
-  }
-}
-\addbibresource{References.bib}
-\begin{document}
-\maketitle
-\section*{Introduction}
-(What needs to be explained? What is the puzzling observation? What is the research question?)
-Has China's aid project changed how the United States conducts it's aid?
-I argue that we should expect the Chinese project to have an effect because the United States has made internet freedom part of it's foreign aid policy, which is perhaps being affected by Chinese aid itself.
-Been a lot of interest recently on the effects of China's BRI (e.g., cite some highly published journals).
+ U.S.
+attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).
 
-Evidence that internet freedom has been a part of U.S.
-foreign policy for the past decade, but no observational evidence that U.S.
-foreign aid (an important part of U.S.
-foreign policy) is related to the internet freedom of recipient countries.
-Further, it is not ex ante obvious why the U.S.
-should rationally be spending money on a relatively abstract concept of internet freedom (literature on how concern for democracy drives U.S.?
-There's a lot of literature on if the US's aid program is self-interested or altruistic, which can leverage here (i.e., given that U.S.
-seems to be self-interested, why?))
-
-I argue that to the extent that U.S.
-aid is related to internet freedom, it is rational for the U.S.
-to expend resources on increasing internet freedom because 1) it secures them access for the future to developing country's markets that could give profitable personal data to U.S.
-Big Tech (there is a possibility that internet is a zero-sum game between China, U.S., where U.S.
-wants to preempt China's takeover of technology markets, i.e.
-technological arms race, access war) and 2) the U.S.
-has an interest in thwarting China's efforts at increasing domestic economic growth through their technology companies.
-
-\section*{Literature Review}
-\subsection*{Determinants of Aid}
-\textbf{Self-Interest versus Altruism}
-
-Hoeffler, Outram: ``Our results indicate that all bilateral donors allocate aid according to their self-interest and recipient need''
-
-Younas: Trade seems to account for donor behavior, along with priorities in human rights and miseries rather than poverty.
-
-Wiseman, Young: Donors base behavior off of ``warm-glow''
-
-National interests versus recipient need (general debate about self-interest or in the interest of recipients).
-
-\section*{Theory}
-It is rational for the U.S. to (condition?) spend foreign aid according to internet freedom levels of recipient because...
-
-\section*{Methodology}
-
-\section*{Complications}
-
-\section*{Conclusion}
-
-\pagebreak
-\printbibliography
-\end{document}
+ DiD?
+Similar levels of internet freedom in the past, then Chinese aid begins to increase, then what happens to U.S.
+aid?
+Or backwards?
+(U.S.
+aid trends and relatively similar (parallel trends) then Chinese aid begins) (OR look at two countries with different internet freedom levels that have received relatively similar levels of US Aid, then see what happens after Chinese aid begins) if no DiD, I can definitely do this as a case study.

### 2022-10-01 Paper.tex
new
Progress

@@
 The United States (U.S.) has been openly concerned about the internet since at least 1997, when President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (\cite{government1997}).
-By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as internet freedom programs cropped up in different facets of the U.S.
-government, such as the Department of State (e.g., \cite{government2021}) and the U.S.
-Agency for Global Media (\cite{government2022d}).
+By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as so-called internet freedom programs cropped up in different facets of the U.S.
+government, such as the Department of State (e.g., \cite{henry2014}; \cite{government2021a}) and the U.S.
+Agency for Global Media (\cite{government2022}).
+
 U.S.
-attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).
+attention to internet freedom has continued to grow (e.g., \cite{government2010}), crescendoing most recently in 2022.
+All in one year, bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S.
+and over 60 other countries to combat global ``digital authoritarianism'' (\cite{government2022b}), and an unprecedented, U.S.-led G7 spending plan was unveiled to improve physical and digital infrastructure abroad (\cite{government2022c}).
 
-Today internet freedom has seemingly reached a crescendo in U.S.
-foreign policy.
-Bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S.
-and over 60 other countries to combat ``digital authoritarianism'' (\cite{government2022b}), and an unprecedented, U.S.-led G7 spending plan was unveiled to improve physical and digital infrastructure abroad (\cite{government2022c})- all in 2022.
-
-Along the way, the People's Republic of China (China) and the Russian Federation (Russia) have been named the antagonists to global internet freedom and the explicit motivations behind the U.S.'s focus on the subject.
+Along the way, the People's Republic of China (China) and the Russian Federation (Russia) have been named the antagonists to global internet freedom and the explicit motivations behind the U.S.'s focus on the subject (see \cite{government2010}, \cite{government2022b}).
 During this intensifying period of U.S.
-apprehension of digital authoritarianism, China launched it's Belt and Road Initiative (BRI) in 2013 and the accompanying Digital Silk Road Initiative (DSR) in 2015.
+apprehension of apparent digital authoritarianism, China launched it's Belt and Road Initiative (BRI) in 2013 and the accompanying Digital Silk Road Initiative (DSR) in 2015.
 Collectively, these projects have catalyzed over one trillion U.S.
@@
 
-Ample academic work has recently attempted to outline the effects of China's aid program (e.g., \cite{gehring2022}; \cite{dreher2018}; \cite{blair2021}).
+Ample academic work has recently attempted to outline the effects of China's aid program (e.g., \cite{blair2021}; \cite{eichenauer2021}; \cite{dreher2018}).
 Yet, little research has explored the potential effects of China's spending on U.S.
@@
 foreign aid, an important aspect of foreign policy, can be understood as a response to China's aid ambitions.
+
+Literature shows that foreign aid is typically a self-interested tool (e.g., \cite{hoeffler2011}) and that U.S.
+aid can be effective in stimulating liberal democratic values in a contemporary environment saturated with Chinese aid (\cite{blair2022}).
+Given these findings, it seems likely that a view of U.S.
+foreign aid over time would reflect the historical policy importance of internet freedom and the U.S.'s desire to oppose China's growing influence and authoritarian model of the internet.
 Specifically, I hypothesize that U.S.
@@
 
-\section*{Literature Review}
-Most foreign aid allocated to self-interest and not recipient need (\cite{hoeffler2011})
+Should this relationship be borne out in the data, future work would have to precisely explain how the U.S.
+benefits from higher global levels of internet freedom if foreign aid is indeed a self-interested tool.
+If no relationship is apparent, however, it must similarly be explained why the U.S.
+is failing to reconcile their foreign policy position on internet freedom with their intention to compete with China's expanding aid project since there is evidence that foreign aid can be effective in promoting liberal democratic values.
 
-Aid ineffectiveness in achieving economic growth or promoting democratic institutions (Chong, Gradstein, Calderon)
+I proceed by highlighting the relevant literature on foreign aid before moving on to the operationalization and testing of the hypothesis.
+In the concluding remarks, I offer tentative theories for various anticipated results.
 
-Aid from U.S. successful in promoting democracy (\cite{blair2022})
+\section*{Literature Review} %improve literature review: both the number and quality of discussion on citations; it is imperative that I can claim that U.S.
+interest in internet freedom makes sense in terms 1) self-interest and 2) effectiveness in the current aid climate.
+Also, maybe contain a few citations about how the U.S.
+as a specific case creates foreign aid preferences (e.g., Milner and folks)
+\subsection*{Determinants of Foreign Aid: Self-Interest or Altruism?}
+Scholars have long debated if donors of foreign aid are motivated by altruism or self-interest.
+The received wisdom on foreign aid states that donors often give aid according to their strategic interests (e.g., \cite{mckinlay1977}), and economic trade has been shown to be a predictor of donor aid.
+Specifically, donors seem to distribute aid amongst recipients that import goods that donors have a comparative advantage in producing (\cite{younas2008}).
 
-Aid from U.S.
-tends to be more effective in countries that receive no substantial support from China (so we should see funding directed towards countries lacking Chinese aid already)(Dreher et al 2021 AER)
+However, some studies have implied that altruism can be influential as well.
+France and Japan, for example, have displayed patterns of spending associated with past colonial ties and UN votes, respectively (\cite{alesina2000}).
+Recipient humanitarian need is also thought to be important in determining aid flows and that, despite inefficiencies in allocation, aid still results in substantial poverty reduction (\cite{collier2002}).
+However, more recent work on recipient need has seemed to conclude that it is largely dominated by donor self-interest, although donor behavior is still poorly understood (\cite{hoeffler2011}.
 
-Self-interest (Alesina and Dollar)
+There is also seeming variation amongst donor countries depending on if they are a part of the Development Assistance Committee (DAC).
+More traditional donors that operate with DAC are shown to care more for recipient need while nonDAC donors allocate aid more so with regard to self-interest (\cite{dreher2011}).
+Generally, though, DAC aid providers care less for recipient need than previously thought while nonDAC members are less self-interested than expected by previous literature.
 
-Chinese aid associated with increase in authoritarian values; World bank aid associated with democracy (Gehring, Kaplan, and Wong)
+\subsection*{Foreign Aid: An Effective Tool?}
+Regardless of the motivations behind foreign aid, there has been a significant amount of disagreement over whether foreign aid is effective in achieving its objectives.
+Aid, for example, has been found to be ineffective in altering levels of inequality or poverty in recipient countries, even if those countries have relatively strong institutions (\cite{chong2009a}).
+The lack of intentionality behind aid allocation is thought to contribute to its ineffectiveness, with donors not necessarily interested in maximizing aid effectiveness on poverty reduction or humanitarian need (\cite{alesina2000}, \cite{collier2002}).
 
-aid increases democracy (Dietrich, Wright)
+There is, however, modest evidence that economic aid contributes to democratic consolidation in Africa (\cite{dietrich2015a}).
 
-Commercial self-interest present but possibly overblown (Dreher, Nunnenkamp, Thiele)
+Interestingly, evidence for the effectiveness of aid appears to have increased with China's monumental entrance into aid provision.
+For example, when comparing the effects of Chinese and U.S.
+aid in Africa, U.S.
+aid is relatively successful in promoting liberal democratic values amongst Afrobarometer respondents (\cite{blair2022}).
+Chinese aid, by contrast, largely has null or even positive effects on democratic values as well.
 
-Bilateral aid more likely to be conditioned on things and less impartial, implying that U.S. should use it more if it is really concerned with internet freedom changes (Christensen, Homer, Nielson)
+Further research on aid's impact on democratic preferences in Africa after the rise of Chinese aid finds that aid from the World Bank similarly increased liberal democratic attitudes (\cite{gehring2022}).
+World Bank aid is even associated with improved stability, whereas Chinese aid is shown to have relatively smaller effects on stability and a positive relationship to autocratic acceptance.
 
-Trade benefits drive aid (Younas)
+There is a substantial amount of nuance to aid effectiveness, though.
+Aid from China is associated with positive economic growth in recipient countries while U.S.
+aid is relatively more effective in countries not currently receiving Chinese aid (\cite{dreher2021}).
 
-Political economic factors can shape preferences for foreign aid in U.S. (Milner, Tingley)
+Other work details how bilateral aid is more likely to be conditioned on levels of corruption and thus more effective than multilateral aid (\cite{christensen2011}).
 
 \section*{Methodology}
-Dependent variable: US Aid
+Following the logic and stated hypothesis in the introduction, the dependent variable is U.S.
+foreign aid to countries in Africa while the main explanatory variables are recipient country internet freedom levels and Chinese foreign aid to Africa.
+The time-span of the data range from 2000-2017, with the data described as follows.
 
-Main explanatory variables: Chinese Aid, Internet Freedom
+U.S.
+foreign aid is measured using disbursement data from the U.S.
+government (\cite{government2022e}), which reports foreign aid on a fiscal-year basis.
+The measurement for Chinese foreign aid stems from the College of William \& Mary (\cite{custer2021}) which quantifies official financial and in-kind commitments made by China.
+Internet freedom data is provided by the Digital Society Project (DSP, \cite{mechkova2022}), being collected through expert surveys and estimated with methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}).
+The DSP dataset is potentially superior for measuring the concept of internet freedom compared to the alternative, Freedom of the Net (FOTN, \cite{house2022}).
+FOTN features frequent missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.
 
-Controls (according to Cinelli et al.): GDP per capita (recipient need), Internet Usership (Internet development indicators), Political Freedom (possible confounder)
+According to the most recent advice for determining good controls (\cite{cinelli2022}), I include GDP per capita, an indicator for internet development and political freedom as controls.
+GDP per capita is assessed in terms of 2021 U.S.
+dollars as is supplied by the World Bank (\cite{bank2022}).
+The indicator for internet development, internet usage by country, is calculated as a percentage and is reported by the International Telecommunication Union (ITU, \cite{itu2022}).
+Lastly, the chosen measure for political freedom comes from V-Dem (\cite{coppedge2022a}).
 
-Chinese data set (Custer et al. 2021; Dreher et al. 2022 (Banking on Beijing))
+OLS with country fixed effects
 
-If concerned about simultaneity between US aid and Internet Freedom, maybe use instrument related to Internet Freedom but not to US aid.
+IV?
 
-DiD? With two countries that have similar levels of Internet Freedom and Chinese aid (parallel trends) and then both experience increases in Chinese aid?
-\section*{Complications}
+Case study comparing U.S. aid changes to countries that had similar internet freedom levels but experienced different influxes of Chinese aid.
+
+\section*{Limitations}
+Definition of internet freedom (difference between DSP, U.S.)
+
+Endogeneity.
 
 \section*{Conclusion}
-(Speculate about why US might be rationally interested in increasing internet freedom) Given the literature on the self-interestedness of donors, if U.S.
-aid is changing in response to Chinese aid, what might be a rational motivation?
-Basically: aid effectually increases democracy, democracy is good for business (American Big Tech).
+Restate important parts of the introduction.
 
+Speculate on theoretical reasons why U.S.
+aid might or might not be related to internet freedom and Chinese foreign aid.
+If positively related, say that internet freedom is good for capitalist business.
+If null, perhaps data doesn't capture recent trends as focus has increased.
+Also, perhaps government decision making slow to react (Congress) or unwilling for domestic reasons (Milner stuff, for example).
+
+\nocite{mechkova2022a}
+\nocite{pemstein2022}
+\nocite{coppedge2022}
+\nocite{coppedge2022a}
+\nocite{coppedge2022b}
+\nocite{coppedge2022c}
+\nocite{coppedge2022d}
 \pagebreak

### 2022-10-02 Paper.tex
Five-Pager

@@
 \usepackage[utf8]{inputenc}
-\title{\vspace{-2cm}Digital Authoritarianism and United States Foreign Aid\vspace{-0.5cm}}
+\title{\vspace{-2.75cm}Digital Authoritarianism and United States Foreign Aid\vspace{-0.5cm}}
 \author{Nicholas Ray}
-\date{\vspace{-0.5cm}September 2 2022\vspace{-1cm}}
+\date{\vspace{-0.30cm}October 2 2022\vspace{-1cm}}
 \usepackage[margin=1in]{geometry}
@@
 The United States (U.S.) has been openly concerned about the internet since at least 1997, when President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (\cite{government1997}).
-By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as so-called internet freedom programs cropped up in different facets of the U.S.
+By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as so-called ``internet freedom'' programs cropped up in different facets of the U.S.
 government, such as the Department of State (e.g., \cite{henry2014}; \cite{government2021a}) and the U.S.
@@
 Given these findings, it seems likely that a view of U.S.
-foreign aid over time would reflect the historical policy importance of internet freedom and the U.S.'s desire to oppose China's growing influence and authoritarian model of the internet.
+foreign aid over time would reflect the historical policy importance of internet freedom and the U.S.'s desire to oppose both China's growing influence and their authoritarian model of the internet.
 Specifically, I hypothesize that U.S.
@@
 
+Understanding the U.S.'s actualization of internet preferences has important implications not only for the foreign aid literature, but also for international relations and comparative scholars more broadly.
+Firstly, any result that indicated the importance of internet considerations in inter-state behavior would be the first of its kind to my knowledge, and would signify a potentially novel dimension along which states interact.
+Secondly, this paper would help inform the dynamics of the U.S.
+and China's ongoing contest for economic and political influence by highlighting another issue area by which they compete.
+Lastly, this work would raise questions about the generalizability of state concern over the internet, which could encompass studies on conflict and cyberspace, research on democratization and information asymmetries, and political economic work on globalization.
+
 I proceed by highlighting the relevant literature on foreign aid before moving on to the operationalization and testing of the hypothesis.
@@
 
-\section*{Literature Review} %improve literature review: both the number and quality of discussion on citations; it is imperative that I can claim that U.S.
-interest in internet freedom makes sense in terms 1) self-interest and 2) effectiveness in the current aid climate.
-Also, maybe contain a few citations about how the U.S.
-as a specific case creates foreign aid preferences (e.g., Milner and folks)
-\subsection*{Determinants of Foreign Aid: Self-Interest or Altruism?}
+\section*{Literature Review}
+\subsection*{Altruistic versus Self-Interested Aid}
 Scholars have long debated if donors of foreign aid are motivated by altruism or self-interest.
-The received wisdom on foreign aid states that donors often give aid according to their strategic interests (e.g., \cite{mckinlay1977}), and economic trade has been shown to be a predictor of donor aid.
+The received wisdom on foreign aid states that donors often give aid according to their strategic concerns (e.g., \cite{mckinlay1977}; \cite{alesina2000}), and economic trade has more recently been shown to be a predictor of donor aid.
 Specifically, donors seem to distribute aid amongst recipients that import goods that donors have a comparative advantage in producing (\cite{younas2008}).
 
-However, some studies have implied that altruism can be influential as well.
-France and Japan, for example, have displayed patterns of spending associated with past colonial ties and UN votes, respectively (\cite{alesina2000}).
-Recipient humanitarian need is also thought to be important in determining aid flows and that, despite inefficiencies in allocation, aid still results in substantial poverty reduction (\cite{collier2002}).
-However, more recent work on recipient need has seemed to conclude that it is largely dominated by donor self-interest, although donor behavior is still poorly understood (\cite{hoeffler2011}.
+However, some studies have implied altruism is also an important factor in determining aid flows, especially for private aid and non-governmental organizations (\cite{buthe2012}).
+Other scholars have analyzed the potential for aid to reduce poverty, concluding that poverty reduction is occurring but in an inefficient manner (\cite{collier2002}).
+The consensus, though, still seems to be that recipient humanitarian need is largely dominated by donor self-interest (e.g., \cite{hoeffler2011}).
 
-There is also seeming variation amongst donor countries depending on if they are a part of the Development Assistance Committee (DAC).
-More traditional donors that operate with DAC are shown to care more for recipient need while nonDAC donors allocate aid more so with regard to self-interest (\cite{dreher2011}).
-Generally, though, DAC aid providers care less for recipient need than previously thought while nonDAC members are less self-interested than expected by previous literature.
+The balance between strategic interests and altruism has additionally been conceived of along the lines of donor maturity, with more established donors potentially distributing aid in more or less altruistic ways than ``new'' donors (\cite{dreher2011}).
+Generally, though, it is found that traditional aid providers care less for recipient need than previously thought while relatively young donors are less self-interested than expected by previous literature.
 
 \subsection*{Foreign Aid: An Effective Tool?}
-Regardless of the motivations behind foreign aid, there has been a significant amount of disagreement over whether foreign aid is effective in achieving its objectives.
-Aid, for example, has been found to be ineffective in altering levels of inequality or poverty in recipient countries, even if those countries have relatively strong institutions (\cite{chong2009a}).
-The lack of intentionality behind aid allocation is thought to contribute to its ineffectiveness, with donors not necessarily interested in maximizing aid effectiveness on poverty reduction or humanitarian need (\cite{alesina2000}, \cite{collier2002}).
+Regardless of the exact source of motivation behind foreign aid, there has been a significant amount of disagreement over whether foreign aid is effective in achieving any objectives.
+In some cases, for example, it appears that the lack of discernible and consistent motivations is the very reason why aid can be found to be ineffective (\cite{collier2002}).
+Aid has also been found to be ineffective in altering levels of inequality or poverty in recipient countries, even if those countries have relatively strong institutions (\cite{chong2009a}).
 
-There is, however, modest evidence that economic aid contributes to democratic consolidation in Africa (\cite{dietrich2015a}).
+There is, however, modest evidence that economic aid effectively contributes to democratic consolidation in Africa (\cite{dietrich2015a}).
+Effectiveness can vary by certain characteristics of aid as well, with bilateral aid being more likely to be conditioned on a recipient country's level of corruption.
+Multilateral aid, on the other hand, is generally more impartial and unconditional, resulting in diminished effectiveness when compared to bilateral aid (\cite{christensen2011}).
 
-Interestingly, evidence for the effectiveness of aid appears to have increased with China's monumental entrance into aid provision.
+Interestingly, evidence for the effectiveness of aid appears to have strengthened with China's monumental entrance into the aid arena.
 For example, when comparing the effects of Chinese and U.S.
 aid in Africa, U.S.
-aid is relatively successful in promoting liberal democratic values amongst Afrobarometer respondents (\cite{blair2022}).
-Chinese aid, by contrast, largely has null or even positive effects on democratic values as well.
+aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents (\cite{blair2022}).
+Chinese aid, by contrast, largely has null or even similarly positive effects on democratic values.
+In a separate study relating World Bank and Chinese aid, foreign aid from the World Bank is associated with the same liberal democratic values and even improved country stability.
+A feeling of acceptance towards autocratic leadership is found to be connected to Chinese aid (\cite{gehring2022}).
 
-Further research on aid's impact on democratic preferences in Africa after the rise of Chinese aid finds that aid from the World Bank similarly increased liberal democratic attitudes (\cite{gehring2022}).
-World Bank aid is even associated with improved stability, whereas Chinese aid is shown to have relatively smaller effects on stability and a positive relationship to autocratic acceptance.
+Yet there is a substantial amount of nuance to aid effectiveness.
+Aid from China is definitively linked with positive economic growth in recipient countries, for example, while U.S.
+aid is most effective in countries not currently receiving large amounts of Chinese aid (\cite{dreher2021}).
 
-There is a substantial amount of nuance to aid effectiveness, though.
-Aid from China is associated with positive economic growth in recipient countries while U.S.
-aid is relatively more effective in countries not currently receiving Chinese aid (\cite{dreher2021}).
+Thus, it can be gathered from previous literature that donor behavior, including that of the U.S., should be expected to be a product of strategic considerations.
+Furthermore, while not always effective in securing these considerations, it is plausible that the current atmosphere of aid-giving in Africa represents a scenario in which aid tied to liberal democratic ideals can be effectual.
 
-Other work details how bilateral aid is more likely to be conditioned on levels of corruption and thus more effective than multilateral aid (\cite{christensen2011}).
+\section*{Empirics}
+\subsection*{Data and Variables}
+In approaching an analysis of the relationship between foreign aid and internet freedom, it perhaps makes most sense to firstly define the main explanatory variable of internet freedom.
+The Digital Society Project (DSP) aims to identify ``how people use social media as a political tool and to explore how political institutions and social media use interact,'' providing data from expert surveys that cover a range of topics such as censorship, disinformation, and monitoring (\cite{mechkova2022}).
+The DSP employs methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}) to try and estimate differences in expert opinion and knowledge.
 
-\section*{Methodology}
-Following the logic and stated hypothesis in the introduction, the dependent variable is U.S.
-foreign aid to countries in Africa while the main explanatory variables are recipient country internet freedom levels and Chinese foreign aid to Africa.
-The time-span of the data range from 2000-2017, with the data described as follows.
-
-U.S.
-foreign aid is measured using disbursement data from the U.S.
-government (\cite{government2022e}), which reports foreign aid on a fiscal-year basis.
-The measurement for Chinese foreign aid stems from the College of William \& Mary (\cite{custer2021}) which quantifies official financial and in-kind commitments made by China.
-Internet freedom data is provided by the Digital Society Project (DSP, \cite{mechkova2022}), being collected through expert surveys and estimated with methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}).
-The DSP dataset is potentially superior for measuring the concept of internet freedom compared to the alternative, Freedom of the Net (FOTN, \cite{house2022}).
+Data from the DSP is potentially superior for measuring the concept of internet freedom compared to the primary alternative, Freedom on the Net (FOTN, \cite{house2022}).
 FOTN features frequent missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.
 
-According to the most recent advice for determining good controls (\cite{cinelli2022}), I include GDP per capita, an indicator for internet development and political freedom as controls.
-GDP per capita is assessed in terms of 2021 U.S.
-dollars as is supplied by the World Bank (\cite{bank2022}).
-The indicator for internet development, internet usage by country, is calculated as a percentage and is reported by the International Telecommunication Union (ITU, \cite{itu2022}).
+The other main explanatory variable is Chinese foreign aid, stemming from the College of William \& Mary's AidData research lab (\cite{custer2021}).
+This is the one of the only datasets on Chinese aid and is certainly the largest and most specific.
+Aid is quantified as official financial and in-kind commitments and is measured in constant 2017 USD.
+
+U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{government2022e}). It is measured in constant 2022 USD.
+
+The starting year for the holistic dataset is 2000, defined by the lower bound of data availability for internet freedom, and 2017 is the ending year per data availability for Chinese foreign aid.
+There is some evidence of left-truncation or missing data in the information on internet freedom, but otherwise all African countries are accounted for.
+
+\subsection*{Methodology}
+To estimate the association between U.S.
+foreign aid and the main explanatory variables, I plan to use ordinary least squares (OLS) with country fixed-effects.
+Fixed-effects at the country level help account for any underlying similarity between countries that could induce spuriousness.
+
+In addition to including fixed-effects, there are multiple controls that seem appropriate to incorporate.
+According to the most recent advice for determining good controls (\cite{cinelli2022}), I include gross domestic product (GDP) per capita, an indicator for internet development, and political freedom as controls.
+Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.
+
+GDP per capita is assessed in constant 2021 USD and as is supplied by the World Bank (\cite{bank2022}).
+The indicator for internet development, internet usage by country, is calculated as a percentage and is provided by the International Telecommunication Union (\cite{itu2022}).
 Lastly, the chosen measure for political freedom comes from V-Dem (\cite{coppedge2022a}).
 
-OLS with country fixed effects
+\section*{Limitations}
+There are at least a couple important limitations to be clear about.
+Firstly, there is undoubtedly some measurement error in the quantification of internet freedom brought about due to possible differences in the idea of internet freedom employed by the U.S.
+and DSP.
+In other words, even if the U.S.'s conception of internet freedom was constant over time, the DSP definition is likely to be somewhat distinct.
 
-IV?
-
-Case study comparing U.S. aid changes to countries that had similar internet freedom levels but experienced different influxes of Chinese aid.
-
-\section*{Limitations}
-Definition of internet freedom (difference between DSP, U.S.)
-
-Endogeneity.
+Secondly, there is potential simultaneity between U.S.
+foreign aid and internet freedom.
+More advanced econometric modeling techniques, such as difference-in-differences (DiD) or instrumental variables regression (IV) may be imperative here.
 
 \section*{Conclusion}
-Restate important parts of the introduction.
+It is possible that there is no meaningful association between U.S.
+foreign aid and internet freedom.
+The lack of association would in itself present a sort of puzzle, though, since the U.S.
+clearly has a foreign policy interest in maintaining internet freedom and a seemingly effective tool of foreign aid to pursue that interest.
+It may take time, however, for the ever-increasing advances in U.S.
+financial commitment to internet freedom to be reflected in spending data.
 
-Speculate on theoretical reasons why U.S.
-aid might or might not be related to internet freedom and Chinese foreign aid.
-If positively related, say that internet freedom is good for capitalist business.
-If null, perhaps data doesn't capture recent trends as focus has increased.
-Also, perhaps government decision making slow to react (Congress) or unwilling for domestic reasons (Milner stuff, for example).
+If there is indeed a relationship, I speculate that the U.S.
+is interested in keeping the internet free because it is good for business.
+There is evidence that China is sharing it's technology, like censorship and monitoring tools, that could allow more authoritarian-leaning developing countries to prevent the use of U.S.
+technology and to strictly allow for Chinese substitutes instead.
+These developing countries would then harness the economic benefits of the internet, like job growth, while constraining potentially negative political externalities such as the free assembly of anti-incumbent online groups.
+China would stand to gain domestic economic growth as it's technology firms expand beyond it's borders, and everybody wins- except the U.S.
 

### 2022-11-14 Extra.tex
Five-Pager
Progress

@@
- U.S.
-attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).
+%----------------------------------------------------------
+%To do
+Case study comparing U.S.
+aid changes to countries that had similar internet freedom levels but experienced different influxes of Chinese aid?
+Need to do case studies of results and examples, in general.
+%----------------------------------------------------------
+%Rough Draft extras
+\section*{Literature Review}
+Foreign aid before BRI (a few sentences with biggest takeaways; self-interest, ineffectiveness)
 
- DiD?
-Similar levels of internet freedom in the past, then Chinese aid begins to increase, then what happens to U.S.
-aid?
-Or backwards?
-(U.S.
-aid trends and relatively similar (parallel trends) then Chinese aid begins) (OR look at two countries with different internet freedom levels that have received relatively similar levels of US Aid, then see what happens after Chinese aid begins) if no DiD, I can definitely do this as a case study.
+BRI (describe, including sharing of technology or suspicion of digital authoritarianism)
+
+Foreign aid literature after BRI (a few paragraphs, highlighting how effectiveness has increased and evidence that donors are adjusting, i.e. lower conditionality for example)
+
+\section*{Theory}
+To explain how other donor countries should be expected to react to China's recent dominance of the international financing arena, I model a simple entry game between China and an unspecified country.
+Once China decides to be heavily involved in foreign finance,
+
+Essentially, the payoff a traditional donor gets from increasing their foreign aid spending given that China has ``entered'' the arena is dependent on their level of competition with China.
+I focus on competition along the axis of internet influence...?
+Why this rather than other axes?
+Does it actually capture economic competition, too, as I would want to argue?
+(Foreign aid is self-interested and seemingly effective in this environment, so it seems reasonable that a policy emphasis on IF would trickle down to aid spending...
+right?
+;Perhaps China spreads technology/domestic companies for economic reasons, which could decrease IF and U.S.
+would oppose to not lose market access (Facebook) and would try to bolster IF)...
+
+Where $\gamma$ is the influence that $Donor\;1$ aid has on internet freedom or technology or the internet, and $\delta$ is the ability for $Donor\;2$ to counter this influence or to influence IF, technology, or the internet themselves.
+Then focus on the detailing these assumptions.
+
+\section*{Conclusion}
+Crazy conclusion: I suspect that the U.S.
+and China are in a technological proxy war.
+China, in an effort to boost its global influence and economic growth, is attempting to build digital infrastructure around the world and encouraging the use of Chinese technology.
+In response, the U.S.
+is bankrolling it's own digital infrastructure projects out of fear that its companies will be locked out of markets that subscribe to Chinese technology.
+The setting of this conflict is mainly developing countries that are open to receiving aid from these two countries.
+The U.S.
+prefers countries to have a liberal view of the internet, thereby facilitating the growth of Western technology companies.
+These developing countries, however, have an incentive to harness the economic benefits of the internet while exercising more control over the potential political costs- therefore, they might welcome the Chinese model.
+The Chinese, meanwhile, don't really care but are just aiming to expand their consumer base, but the Chinese model has illiberal externalities.
+
+%----------------------------------------------------------
+%Old theory ideas (explaining an expected relationship)
+\section*{Theory} 
+U.S. wants to protect American Big Tech markets that are potentially going to be lost in a zero-sum game with Chinese companies (if China successfully shares tech and builds infrastructure).
+
+\begin{align*}
+    H_1a:\; & \text{A pattern should be evident of U.S. money flowing into places where either}\\
+    & \text{Big Tech markets are OR}\\
+    H_1b:\; & \text{Where Chinese influence/money is increasing OR}\\
+    H_2:\; & \text{Where Chinese influence is NOT increasing (U.S. defending markets)}
+\end{align*}
+
+\section*{Alternative Explanations}
+1) The U.S. is not concerned necessarily about Big Tech markets, but instead increasing aid in an attempt to thwart China's goals of economic growth abroad.
+
+\textbf{Implication}
+\begin{align*}
+    H_3:\; & \text{There should be a relatively random pattern of spending or at least}\\
+    & \text{a pattern not associated with Big Tech markets}
+\end{align*}
+
+2) The U.S.
+is not necessarily concerned about market access for American Big Tech or China's goal of increasing economic growth through trade/economic diplomacy, but rather trying to counter Chinese money's influence on digital authoritarianism.
+The U.S.
+would then be trying to share it's own tech and build it's own internet infrastructure to have more control over the recipient's internet freedom levels.
+
+\textbf{Implication}
+\begin{align*}
+    H_4:\; & \text{U.S. spending patters would be associated with increasing}\\
+    & \text{internet freedom levels and potentially concentrated in places that are}\\
+    & \text{not already receiving tech and infrastructure from China}\\
+\end{align*}
+
+9/28
+If U.S.
+foreign aid policy is rational and related to internet freedom, then the U.S.
+must stand to gain from spending money on improving others' internet freedom.
+I argue that the U.S.
+benefits in two main ways from supporting internet freedom: protecting American technology's access to developing markets (possibly being encouraged by China to use Chinese technology instead) while ensuring that China faces resistance in continuing its economic growth (which is it's main goal of this project).
+
+If this is true, then we should observe U.S. foreign aid spending patterns with one set of the following hypotheses:
+
+\begin{align*}
+    H_1a:\; & \text{U.S. aid should be associated with countries with high levels of internet freedom.}\\
+    H_1b:\; & \text{U.S. aid should be associated with countries with low levels of internet freedom.}\\
+    H_2a:\; & \text{U.S. aid should be associated with countries receiving high amounts of Chinese aid.}\\
+    H_2b:\; & \text{U.S. aid should be associated with countries receiving low amounts of Chinese aid.}
+\end{align*}
+
+9/29
+\section*{Theory}
+Depending on how the literature review goes, we can mention our theory of why it is rational for the U.S.
+to engage in promoting internet freedom (aid effectually increases democracy, and democracy is good for business (American Big Tech)).
+
+%----------------------------------------------------------

### 2022-11-14 RoughDraft.tex
Progress

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\title{\vspace{-2.75cm}628\vspace{-0.5cm}}
\author{Nicholas Ray}
\date{\vspace{-0.30cm}November 14 2022\vspace{-1cm}}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\doublespacing
\usepackage{footmisc}
\renewcommand{\footnotelayout}{\setstretch{1.75}}
\usepackage{footmisc}
\renewcommand{\footnotesize}{\normalsize}
\usepackage{caption}
\usepackage{tikz}
\usepackage{istgame}
\usepackage[backend=biber, style=authoryear, maxbibnames=99,uniquelist=false]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \clearfield{month}
  \clearfield{day}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{ForeignAidBib.bib}
\begin{document}
\maketitle
\begin{abstract}

\end{abstract}
\section*{Introduction}
Research on foreign aid has seemingly concluded that donors use aid as a self-interested policy tool (e.g., \cite{alesina2000}; \cite{hoeffler2011}) and that its effectiveness greatly varies (e.g., see \cite{christensen2011})\footnote{All data and code for replication purposes can be found on GitHub at https://github.com/nnray/Analysis. In the ``Code'' folder, navigate to ``628.Rmd'' for my R script. My full \LaTeX \;file can be found at https://github.com/nnray/628.}. Contemporary foreign aid work has largely sought to understand how the recent influx of aid from the People's Republic of China (China) has affected these conclusions (see \cite{dreher2022} on China's spending). There is some evidence that traditional, Western donors have updated how they give aid and that aid effectiveness has also changed. For example, albeit loans, the World Bank appears to have reduced the conditions it places on borrowing countries (\cite{hernandez2017}) while aid from the United States (U.S.) seems to have become more effective at stimulating liberal democratic values (\cite{blair2022}).

Thus far, this literature has overlooked a potentially salient factor for how Western donors allocate aid in a world increasingly exposed to Chinese finance: the internet. China is thought to be exporting technology alongside its aid programs in a bid to spread its influence and stimulate domestic economic growth, although countries like the U.S. accuse China of intentionally promoting so-called ``digital authoritarianism'' (Clinton 2011, declaration 2022). While seemingly trivial, this viewpoint of the U.S. reflects the increasing competition between the two countries for technological superiority (Hass et al.). Given that Chinese aid has become more prominent and that there is evidence that Western aid practices are responding to this, it's a natural question to ask how much of this response is oriented around technological concerns and the internet.

More pointedly, is the U.S. allocating foreign aid in a way to compete with China technologically? If so, ... And if not, ... 
%Contribution? What would knowing that this conflict is occurring tell you? China and the LIO?
%Interesting because if yes, why is U.S. spending so much on IF if only for some ideological goal (i.e., might not be rational) and if no, why is U.S. failing to take opportunity (not rational)?

I proceed as follows...
%In the next section, I summarize my argument before delving into the assumptions underpinning my theory. Followed by results, discussion, conclusion.

\section*{Theory}
\subsection*{Model}
I argue that it is rational for the U.S. to consider the internet when distributing foreign aid and, specifically, that a recipient's current use of the internet should be a strong predictor for receiving U.S. funds. Consider an entry deterrence game, where China is the first-mover and decides to ``enter'' the foreign aid arena in a particular country by allocating aid or not. The U.S. then moves second, making a similar decision to distribute foreign aid depending on if China has entered or not. The U.S. bases its rationale to contribute foreign aid in a country on the ability of its aid to influence internet governance relative to the ability of China's aid to do the same.

When U.S. foreign aid is net effective at encouraging a liberal model of the internet in a country, it decides to allocate aid. China, observing the U.S.'s expenditure, decides not to distribute foreign aid in that country. However, if the U.S. concludes not to spend aid monies due to its aid being relatively ineffective at influencing a recipient's use of the internet, then China allocates aid and exerts its influence on a recipient's internet usage. The extensive-form of this game is displayed in Figure 1.

\subsubsection*{Figure 1}
\begin{istgame}[scale=4]
\xtdistance{5mm}{15mm}
\istroot(0)(0,0){\textit{China}}
\istb{\neg Spend}[al]
\istb{Spend}[ar]{0\leq\gamma,c,\delta\leq1}[below,yshift=25mm,xshift=25mm]
\endist
\xtdistance{5mm}{10mm}
\istroot(1)(0-1)<above left>{\textit{U.S.}}
\istb{\neg Spend}[l]{(\gamma,\;\delta)}
\istb{Spend}[r]{(\gamma,\;\frac{1}{\delta}-c_2)}
\endist
\istroot(2)(0-2)<above right>{\textit{U.S.}}
\istb{\neg Spend}[l]{(\frac{1}{\gamma}-c_1,\;\delta)}
\istb{Spend}[r]{(\frac{\delta}{\gamma}-c_1,\;\frac{\gamma}{\delta}-c_2)}
\endist
\xtdistance{10mm}{10mm}
\end{istgame}

\subsection*{Technical Assumptions}
Ample explanation is in order. Firstly, there is a host of technical restrictions inherent to my argument. Other determinants of foreign aid, for example, are not considered here. Thus, it is assumed that the U.S. and China spend aid in consideration of internet influence $ceteris\;paribus$, or holding other foreign aid avenues constant.

These decisions are also stipulated as dichotomous, where China or the U.S. simply agree to spend foreign aid or not. In reality, foreign aid spending is clearly continuous in nature.

Further, this is a full information model. There is no uncertainty over how the U.S. or China spends foreign aid money.

Lastly, but not exhaustively, I assume that influence over internet politics for the U.S. is inversely related to Chinese influence and vice versa.

\subsection*{Substantive Assumptions}
Secondly, and most critically, there are a number of substantive assumptions invoked in this theoretical exposition that must be detailed. For one, I am assuming that the U.S. has a preference over internet governance around the world and, additionally, that foreign aid is an effective instrument for the U.S. to actualize this preference. Somewhat symmetrically, I assume that Chinese aid has an effect on a recipient's internet model and that this effect is negative or illiberal. I will take these assumptions in turn, connecting these behavioral claims to the established literature where possible.

Beginning with the U.S., there is suggestive evidence that the U.S. has a preference over how other countries use the internet. Most recently, bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S. and over 60 other countries to combat global ``digital authoritarianism'' (\cite{government2022b}). These concerns stretch back to at least 2010, when then-Secretary of State Hillary Clinton (\cite{government2010}). Even before then, the U.S. Department of State was... (\cite{government2021a}).

If one accepts that the U.S.'s apparent foreign policy position on the internet coincides with more traditional accounts of U.S. preferences for democracy world-wide, then there is indirect evidence that foreign aid is an effective tool for influencing internet politics in recipient countries. Most pertinent is \cite{blair2022}, who show how...when comparing the effects of Chinese and U.S. aid in Africa, U.S. aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents. However, evidence that U.S. aid is most effective in countries not currently receiving large amounts of Chinese aid (\cite{dreher2021}).

Turning to China, there is reason to believe that Chinese aid could influence a recipient's political use of the internet in an illiberal manner. China launched the Belt and Road Initiative (BRI) in 2013 and the Digital Silk Road (DSR) in 2015, both aimed at enlarging China's global influence while strengthening economic growth.
%In actuality, China's goals and abilities are very complex
(Shen 2018) non-unitary actor, complexities of state vs private goals
(Tugendhat) the fact that digital funding seems to have decreased after official launch of DSR
%but what is most pertinent to this study is that the West see's it more simply:
(Hillman 2021) example of Western anxiety about China digitzing whole world
(Triolo and Greene) another example but pretty solid info
%And, some evidence that Chinese aid does have autocratic effect
A feeling of acceptance towards autocratic leadership is found to be connected to Chinese aid (\cite{gehring2022}).

%Talk about BRI and DSR together lead to China having developed a lot of digital infrastructure in developing countries (how could this negatively affect internet) and how China is sharing technology to increase state capacity over the internet (and make clear how this negatively affects internet). Shen 2018. Tugendhat 2021. IISS webpage. https://www.chinausfocus.com/finance-economy/chinas-digital-silk-road-a-pathway-to-post-pandemic-economic-recovery. Hillman 2021. Naughton 2020. Triolo Carnegie webpage

\subsection*{Implications}
Returning to my theory, the model simplistically states that the U.S. and China should allocate foreign aid when aid is relatively effective at influencing a recipient's political model of the internet. The literature implies that U.S. aid is effective when Chinese aid is low or not present. Since it is assumed that Chinese aid has a negative effect on IF, this further implies that U.S. aid should be being spend in countries with higher IF. These two implications lead to the following hypotheses:
\begin{align*}
    &\text{``Interstate Competition Hypothesis''}\\
    H_{1_a}&:\;\text{U.S. aid inversely related to Chinese aid}\\
    &\text{``Trickle Down Hypothesis''}\\
    H_{2_a}&:\;\text{U.S. aid directly related to IF}\\
\end{align*}
However, it is also hypothetically possible that U.S. would try to compete with Chinese aid by directly countering it in countries where the Chinese are active. Similarly, it could make sense for the U.S. to be focusing on improving IF in countries with already low levels, as its aid could be more effective here... This opposing logic generates the following hypotheses:
\begin{align*}
    &\text{``Intrastate Competition Hypothesis''}\\
    H_{1_b}&:\;\text{U.S. aid directly related to Chinese aid}\\
    &\text{``Bottom Up Hypothesis''}\\
    H_{2_b}&:\;\text{U.S. aid inversely related to IF}\\
\end{align*}

\section*{Empirics}
\subsection*{Setting}
I test these hypotheses by evaluating the relationship between U.S. aid, Chinese aid, and the internet across 44 African countries between 2010 and 2017. The full set of countries includes: Algeria, Angola, Benin, Botswana, Burundi, Cabo Verde, Cameroon, Central African Republic, Chad, Comoros, Republic of the Congo, Cote d'Ivoire, Democratic Republic of the Congo, Djibouti, Egypt, Equatorial Guinea, Ethiopia, Gabon, Gambia, Ghana, Guinea, Guinea-Bissau, Kenya, Lesotho, Liberia, Madagascar, Malawi, Mali, Mauritania, Mauritius, Morocco, Mozambique, Namibia, Niger, Nigeria, Rwanda, Senegal, Sierra Leone, South Africa, Togo, Tunisia, Uganda, Zambia, and Zimbabwe.

I chose to focus on Africa during this time period for three main reasons. Firstly, Africa has been a major area of focus for China finance projects, with the Council on Foreign Relations claiming that ``China already provides more financing for information and communications technology than all multilateral agencies and leading democracies combined do across the continent'' (CFR). Secondly, data for Chinese spending ends in 2017. Thirdly, while 2010 is indeed an arbitrary starting point, both Chinese aid and the development of the internet diminish greatly before this period.

\subsection*{Data and Variables}
%internet freedom
In operationalizing my theory, I measure the political use of the internet as ``internet freedom'' from the Digital Society Project (DSP). The DSP aims to identify ``how people use social media as a political tool and to explore how political institutions and social media use interact,'' providing data from expert surveys that cover a range of topics such as censorship, disinformation, and monitoring (\cite{mechkova2022}). The DSP employs methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}) to try and estimate differences in expert opinion and knowledge.

The DSP is one of two indices measuring a political concept of internet usage and data from the DSP is potentially superior for measuring the concept of internet freedom compared to the primary alternative, Freedom on the Net (FOTN, \cite{house2022}). FOTN features frequent missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.

%Chinese aid
The other main explanatory variable is Chinese foreign aid, stemming from the College of William \& Mary's AidData research lab (\cite{custer2021}). This is the one of the only datasets on Chinese aid and is certainly the largest and most specific. Aid is quantified as official financial and in-kind commitments and is measured in constant 2017 USD.

%U.S. aid
U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{government2022e}). It is measured in constant 2022 USD.

%controls
I include gross domestic product (GDP) per capita, an indicator for internet development, and political freedom as controls. Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.

GDP per capita is assessed in constant 2021 USD and as is supplied by the World Bank (\cite{bank2022}). The indicator for internet development, internet usage by country, is calculated as a percentage and is provided by the International Telecommunication Union (\cite{itu2022}). Lastly, the chosen measure for political freedom comes from V-Dem (\cite{coppedge2022a}).

%Missing data
Of the 10 African countries I am missing, the DSP is missing values for Somalia, Eritrea, Sudan, and South Sudan, which might make sense given that Somalia and South Sudan were both at civil war during the observation period while Sudan and Eritrea have experienced instability after previous wars. Similarly, Libya was missing data from the ITU following the Western invasion of Libya in 2011. However, it is unclear why Sao Tome and Principe and Seychelles are missing Polity data, why Burkina Faso and Eswatani are missing Chinese aid data, or why Tanzania is missing GDP data.

\pagebreak
Descriptive statistics can be seen in Table 1.

\subsubsection*{Table 1}
\begin{figure}[htbp]
    \includegraphics[scale=1.1]{628table2.png}
\end{figure}

A correlation matrix of the variables included in my regressions is given in Figure 2. As can be seen, there are a couple relatively strong relationships between variables. The variable for regime type, ``Polity 2,'' has a correlation of 0.62 with ``Internet Freedom'' while ``Logged GDP per Capita'' has correlation of 0.65 with the variable measuring internet development, ``Percent Using Internet.'' While possible, I do not think that these correlations are high enough to warrant concerns over multicollinearity or the lack of linear independency in the data. Interestingly, ``Logged U.S. Aid'' appears somewhat negatively correlated with ``Logged GDP per Capita'' and ``Percent Using Internet'' is somewhat positively correlated to ``Internet Freedom.''

\pagebreak
\subsubsection*{Figure 2}
\begin{figure}[htbp]
    \includegraphics[scale=0.7]{628plot3.png}
\end{figure}

\subsection*{Estimation}
I chose to use ordinary least squares (OLS) for my estimator while employing two-way fixed effects at the country and year level. In choosing fixed effects, I used non-nested model comparison tests that suggested fixed effects to be superior to both no effects and random effects. After testing for heteroskedasticity, I included country-clustered standard errors.

\pagebreak
\section*{Results}
Table 2 displays regression results from a model with and without an interaction between internet freedom and logged Chinese foreign aid.

\subsection*{Table 2}
\begin{table}[!htbp] \centering 
  \caption*{Logged U.S. Aid and Covariates with Two-way Fixed Effects and Clustered SE's on Country} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{Logged U.S. Aid} \\ 
\cline{2-3} 
 & Model 1 & Model 2 \\ 
\hline \\[-1.8ex] 
 Internet Freedom & 0.021$^{***}$ & 0.021$^{***}$ \\ 
  & (0.008) & (0.008) \\ 
  & & \\ 
 Logged Chinese Aid & 0.006$^{*}$ & 0.006$^{*}$ \\ 
  & (0.003) & (0.003) \\ 
  & & \\ 
 Logged GDP per Capita & 0.045 & 0.045 \\ 
  & (0.471) & (0.472) \\ 
  & & \\ 
 Percent Using Internet & $-$0.013 & $-$0.013 \\ 
  & (0.009) & (0.009) \\ 
  & & \\ 
 Polity 2 & 0.004 & 0.004 \\ 
  & (0.031) & (0.031) \\ 
  & & \\ 
 Internet Freedom $\cdot$ Logged Chinese Aid &  & 0.00000 \\ 
  &  & (0.0002) \\ 
  & & \\ 
\hline \\[-1.8ex] 
Observations & 352 & 352 \\ 
R$^{2}$ & 0.038 & 0.038 \\ 
Adjusted R$^{2}$ & $-$0.140 & $-$0.144 \\ 
F Statistic & 2.370$^{**}$ (df = 5; 296) & 1.968$^{*}$ (df = 6; 295) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\pagebreak
Seems to support for $H_{1_b}$ and $H_{2a}$. In other words, seems that the U.S., in response to Chinese foreign aid, is using a intrastate competition strategy in competing with China and trying to increase IF in a trickle down manner.

\pagebreak
Figure 3 shows the marginal effect of internet freedom on logged U.S. foreign aid.

\subsection*{Figure 3}
\begin{figure}[htbp]
    \includegraphics[scale=0.7]{628plot1.png}
\end{figure}

\pagebreak
Figure 4 graphs the marginal effect of logged Chinese foreign aid on logged U.S. foreign aid.

\subsection*{Figure 4}
\begin{figure}[htbp]
    \includegraphics[scale=0.7]{628plot2.png}
\end{figure}
\pagebreak

\section*{Discussion}
Talk about implications of trickle-down, intrastate strategy...

Why U.S. might be interested in IF.

Is U.S. response effective at keeping China out? Probably not, although they are ramping up efforts.

\section*{Conclusion}

\pagebreak
\section*{Appendix}
The entry game described in the theory section (Figure 1) is solved via backwards induction. If $China$ chooses to $Enter$, the $Donor$ chooses to $Spend$ if and only if (iff):

\begin{tabular}{c l}
    $\frac{\pi}{\gamma}-c > -1$ &  Condition for $Donor$ to $Spend$, from the payoffs on $Spend$ and $\neg Spend$, respectively.\\
    $\frac{\pi}{c-1} > \gamma $ & Isolating $\gamma$.\\
\end{tabular}

Where $\gamma \in (0,1]$. As $\gamma$ gets arbitrarily small, $\pi$ goes to infinity.

\nocite{mechkova2022a}
\nocite{pemstein2022}
\nocite{coppedge2022}
\nocite{coppedge2022a}
\nocite{coppedge2022b}
\nocite{coppedge2022c}
\nocite{coppedge2022d}
\pagebreak
\printbibliography
\end{document}

### 2022-11-14 RoughDraft.tex
rough draft

@@
 \usepackage[utf8]{inputenc}
-\title{\vspace{-2.75cm}628\vspace{-0.5cm}}
+\title{\vspace{-2.75cm}Conflating Global Finance with Technological Competition: The U.S.'s Reaction to Chinese Foreign Aid\vspace{-0.5cm}}
 \author{Nicholas Ray}
@@
 \begin{abstract}
+    The U.S.
+and China appear to be at variance on the global finance stage and locked in a technological arms race.
+While recent studies on foreign aid have focused on how China's aid programs have changed the conditionality and effectiveness of Western aid, I instead seek to understand how the U.S.
+may have changed its foreign aid strategy in response to competition from China along the dimension of technology and the internet.
+I argue that it is rational for the U.S.
+to react to Chinese aid by conditioning its aid on a recipient's use of the internet in an effort to combat China's suspected negative influence.
+I find that a measure for internet governance and Chinese aid are statistically significant predictors for U.S.
+foreign aid in a sample of 44 African countries between 2010 and 2017.
+\end{abstract}
 
-\end{abstract}
 \section*{Introduction}
-Research on foreign aid has seemingly concluded that donors use aid as a self-interested policy tool (e.g., \cite{alesina2000}; \cite{hoeffler2011}) and that its effectiveness greatly varies (e.g., see \cite{christensen2011})\footnote{All data and code for replication purposes can be found on GitHub at https://github.com/nnray/Analysis.
+Research on foreign aid has seemingly concluded that donors use aid as a self-interested policy tool (e.g., \cite{alesina2000}; \cite{hoeffler2011}) and that its effectiveness greatly varies (e.g., see \cite{christensen2011}).\footnote{All data and code can be found on GitHub at https://github.com/nnray/Analysis.
 In the ``Code'' folder, navigate to ``628.Rmd'' for my R script.
-My full \LaTeX \;file can be found at https://github.com/nnray/628.}.
-Contemporary foreign aid work has largely sought to understand how the recent influx of aid from the People's Republic of China (China) has affected these conclusions (see \cite{dreher2022} on China's spending).
-There is some evidence that traditional, Western donors have updated how they give aid and that aid effectiveness has also changed.
-For example, albeit loans, the World Bank appears to have reduced the conditions it places on borrowing countries (\cite{hernandez2017}) while aid from the United States (U.S.) seems to have become more effective at stimulating liberal democratic values (\cite{blair2022}).
+My full \LaTeX \;file can be found at https://github.com/nnray/628 under ``RoughDraft.tex''.} Contemporary foreign aid work has largely sought to understand how the recent influx of aid from the People's Republic of China (China) has affected these conclusions.
+There is some evidence that traditional, Western donors have updated how they give aid and that aid effectiveness itself has changed.
+For example, albeit with loans, the World Bank appears to have reduced the conditions it places on borrowing countries (\cite{hernandez2017}) while aid from the United States (U.S.) seems to have become more effective at stimulating liberal democratic values (\cite{blair2022}).
 
@@
 China is thought to be exporting technology alongside its aid programs in a bid to spread its influence and stimulate domestic economic growth, although countries like the U.S.
-accuse China of intentionally promoting so-called ``digital authoritarianism'' (Clinton 2011, declaration 2022).
-While seemingly trivial, this viewpoint of the U.S.
-reflects the increasing competition between the two countries for technological superiority (Hass et al.).
-Given that Chinese aid has become more prominent and that there is evidence that Western aid practices are responding to this, it's a natural question to ask how much of this response is oriented around technological concerns and the internet.
+accuse China of promoting so-called ``digital authoritarianism'' (\cite{u.s.departmentofstate2010}; \cite{u.s.departmentofstate2022}).
+The U.S., by contrast, prefers a politically and economically liberal internet where governments invoke the principle of ``laissez faire.'' While seemingly trivial, this viewpoint of the U.S.
+reflects the increasing competition between the two countries for technological superiority (\cite{hassetal.2021}).
 
-More pointedly, is the U.S. allocating foreign aid in a way to compete with China technologically? If so, ... And if not, ... 
-%Contribution? What would knowing that this conflict is occurring tell you? China and the LIO?
-%Interesting because if yes, why is U.S. spending so much on IF if only for some ideological goal (i.e., might not be rational) and if no, why is U.S. failing to take opportunity (not rational)?
+How then has the U.S.
+reacted to an increase in Chinese aid spending around the world, given that China is also a rival on this separate dimension of technology and the internet?
+More pointedly, is the U.S.
+allocating foreign aid in a way that enables it to compete with China technologically?
+I claim that this makes sense given the foreign policy emphasis the U.S.
+has placed on the world-wide internet and the contemporary effectiveness U.S.
+aid has had on liberal values.
+The U.S., by countering aid programs from China that it suspects might negatively affect recipient's model of the internet, can prevent the erosion of internet liberalism in a country.
 
-I proceed as follows...
-%In the next section, I summarize my argument before delving into the assumptions underpinning my theory. Followed by results, discussion, conclusion.
+If this were true, it would be a contribution to both the foreign aid literature and the burgeoning area of work dedicated to understanding China's influence on the liberal international order (LIO).
+Where current work on the LIO focus on how China may or may not be a threat to the pillars of economic or political liberalism (e.g., \cite{weiss}; \cite{broz}), this article can speak to how China is potentially influencing the LIO's similar upholding of internet liberalism.
+China may undermine this tenet, although unintentionally so as a by-product of sharing its technology for economic reasons.
+
+I proceed as follows.
+In the next section, I present my argument and theoretical reasoning for why the U.S.
+might be expected to alter its foreign aid spending in response to the recent spending by China.
+After developing implications and hypotheses, I describe my empirical strategy and data.
+Results and a discussion of the findings are presented before concluding.
 
@@
 funds.
-Consider an entry deterrence game, where China is the first-mover and decides to ``enter'' the foreign aid arena in a particular country by allocating aid or not.
+To explore the rationality of this alleged behavior, I consider a simple entry deterrence game.
+China is the first-mover and decides to enter or spend foreign aid in a particular country by allocating aid or not.
 The U.S.
 then moves second, making a similar decision to distribute foreign aid depending on if China has entered or not.
-The U.S.
-bases its rationale to contribute foreign aid in a country on the ability of its aid to influence internet governance relative to the ability of China's aid to do the same.
+There are two parameters that decide the payoffs of the game: $\gamma$, the influence that Chinese aid has on a recipient country's internet model and $\delta$, the influence of U.S.
+aid.
+Low values represent high degrees of influence for both.
+Essentially, the strategies of the players are symmetrical, with each country deciding to spend aid if their degree of influence is relatively larger than their rival's.
 
@@
 concludes not to spend aid monies due to its aid being relatively ineffective at influencing a recipient's use of the internet, then China allocates aid and exerts its influence on a recipient's internet usage.
-The extensive-form of this game is displayed in Figure 1.
+The extensive-form of this game is displayed in Figure 1, with more detail given in the appendix.
 
@@
 
-\subsection*{Technical Assumptions}
+\subsection*{Notable Assumptions}
 Ample explanation is in order.
@@
 Other determinants of foreign aid, for example, are not considered here.
-Thus, it is assumed that the U.S.
-and China spend aid in consideration of internet influence $ceteris\;paribus$, or holding other foreign aid avenues constant.
+Thus, any changes in spending should be interpreted as the U.S.
+or China deciding to spend on internet influence or not $ceteris\;paribus$, or holding other foreign aid avenues constant.
 
-These decisions are also stipulated as dichotomous, where China or the U.S. simply agree to spend foreign aid or not. In reality, foreign aid spending is clearly continuous in nature.
+These decisions are also stipulated as dichotomous, where China or the U.S.
+simply agree to spend foreign aid or not.
+In reality, foreign aid spending is clearly continuous in nature and, moreover, typically based on repeated interactions between both other donors and recipients.
+This is also a full information model, with no uncertainty present over players' decisions to spend.
+I further assume that payoffs from the decision to spend or not are zero-sum at the extremes, where an arbitrarily large influence by one country dwarfs the incentives for the other to spend foreign aid.
 
-Further, this is a full information model. There is no uncertainty over how the U.S. or China spends foreign aid money.
-
-Lastly, but not exhaustively, I assume that influence over internet politics for the U.S. is inversely related to Chinese influence and vice versa.
-
-\subsection*{Substantive Assumptions}
 Secondly, and most critically, there are a number of substantive assumptions invoked in this theoretical exposition that must be detailed.
@@
 has a preference over how other countries use the internet.
-Most recently, bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S.
-and over 60 other countries to combat global ``digital authoritarianism'' (\cite{government2022b}).
-These concerns stretch back to at least 2010, when then-Secretary of State Hillary Clinton (\cite{government2010}).
-Even before then, the U.S.
-Department of State was...
-(\cite{government2021a}).
+In 2022, bipartisan legislation was introduced in the Senate to increase funding to so-called ``internet freedom programs'' (\cite{foreignrelationscommittee2022}), which were started by the U.S.
+Department of State in the 2000's to combat censorship and surveillance in other countries (\cite{u.s.departmentofstate2021}; \cite{hanson2012}).
+Also in 2022, a declaration was led by the U.S.
+and signed by over 60 other countries to combat global ``digital authoritarianism'' (\cite{u.s.departmentofstate2022}).
+These concerns stretch back to at least 2010, when then-Secretary of State Hillary Clinton (\cite{u.s.departmentofstate2010}) stated the U.S.
+desired a free internet around the world.
 
@@
 preferences for democracy world-wide, then there is indirect evidence that foreign aid is an effective tool for influencing internet politics in recipient countries.
-Most pertinent is \cite{blair2022}, who show how...when comparing the effects of Chinese and U.S.
-aid in Africa, U.S.
-aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents.
-However, evidence that U.S.
-aid is most effective in countries not currently receiving large amounts of Chinese aid (\cite{dreher2021}).
+For instance, in Africa U.S.
+aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents when compared to Chinese foreign aid \cite{blair2022}.
+There is also evidence that U.S.
+aid is most effective at achieving its goals in countries not currently receiving large amounts of Chinese aid (\cite{dreher2021}).
 
 Turning to China, there is reason to believe that Chinese aid could influence a recipient's political use of the internet in an illiberal manner.
-China launched the Belt and Road Initiative (BRI) in 2013 and the Digital Silk Road (DSR) in 2015, both aimed at enlarging China's global influence while strengthening economic growth.
-%In actuality, China's goals and abilities are very complex
-(Shen 2018) non-unitary actor, complexities of state vs private goals
-(Tugendhat) the fact that digital funding seems to have decreased after official launch of DSR
-%but what is most pertinent to this study is that the West see's it more simply:
-(Hillman 2021) example of Western anxiety about China digitzing whole world
-(Triolo and Greene) another example but pretty solid info
-%And, some evidence that Chinese aid does have autocratic effect
-A feeling of acceptance towards autocratic leadership is found to be connected to Chinese aid (\cite{gehring2022}).
+China launched the Belt and Road Initiative (BRI) in 2013 and the Digital Silk Road (DSR) in 2015, both aimed at enlarging China's global influence while strengthening economic growth (\cite{dreher2022}).
+In doing so, it is alleged that China is sharing technology with recipient states that could improve state capacity and allow recipients to curb free expression, both digitally and physically (\cite{hillman2021}).
+Other work, though, rightly points out that China is a non-unitary actor with private firms that could undermine a state-led technology sharing campaign (\cite{shen2018a}) and that China's technology spending has not changed much immediately following the announcement of the DSR (\cite{tugendhat2021}).
 
-%Talk about BRI and DSR together lead to China having developed a lot of digital infrastructure in developing countries (how could this negatively affect internet) and how China is sharing technology to increase state capacity over the internet (and make clear how this negatively affects internet).
-Shen 2018.
-Tugendhat 2021.
-IISS webpage.
-https://www.chinausfocus.com/finance-economy/chinas-digital-silk-road-a-pathway-to-post-pandemic-economic-recovery.
-Hillman 2021.
-Naughton 2020.
-Triolo Carnegie webpage
+However, it seems most pertinent to this study that the West believes China to be spreading its illiberal model of the internet (e.g., \cite{hillman2021}; \cite{u.s.departmentofstate2022}), or at least that there is a good chance for it to (\cite{triolo2020}).
+Further, there is evidence that Chinese foreign aid encourages feelings of acceptance towards autocratic leadership in African countries (\cite{gehring2022}).
 
@@
 and China should allocate foreign aid when aid is relatively effective at influencing a recipient's political model of the internet.
-The literature implies that U.S.
-aid is effective when Chinese aid is low or not present.
-Since it is assumed that Chinese aid has a negative effect on IF, this further implies that U.S.
-aid should be being spend in countries with higher IF.
-These two implications lead to the following hypotheses:
+Some of the literature implies that U.S.
+aid is effective when Chinese aid is low or not present (\cite{dreher2021}).
+Since it is assumed that Chinese aid has a negative effect on how recipients use the internet, this further implies that U.S.
+aid should be spent in countries with more liberal models of the internet to ensure maximum effectiveness.
+These concurrent implications lead to the following hypotheses:
+
 \begin{align*}
     &\text{``Interstate Competition Hypothesis''}\\
-    H_{1_a}&:\;\text{U.S. aid inversely related to Chinese aid}\\
+    H_{1a}&:\;\text{U.S. aid inversely related to Chinese aid}\\
     &\text{``Trickle Down Hypothesis''}\\
-    H_{2_a}&:\;\text{U.S. aid directly related to IF}\\
+    H_{2a}&:\;\text{U.S. aid directly related to recipient internet model}\\
 \end{align*}
-However, it is also hypothetically possible that U.S.
-would try to compete with Chinese aid by directly countering it in countries where the Chinese are active.
+
+I label $H_{1a}$ the ``Interstate Competition Hypothesis'' because if U.S.
+foreign aid is truly most effective where Chinese aid is not present, this implies that the U.S.
+would respond to an increase in Chinese aid by focusing its spending in countries not currently receiving aid from China.
+China and the U.S.
+would thus be spending in completely separate countries or regions.
+
+$H_{2a}$ is titled the ``Trickle Down Hypothesis'' since the U.S.
+would spend more foreign aid in countries with more liberal internet models, implying that the U.S.'s preference for a globally liberal internet model is manifested by it focusing on bolstering countries with already relatively liberal models.
+
+However, it is also hypothetically possible that the U.S.
+would try to compete with Chinese aid by directly countering it in countries where the Chinese are active and generating intrastate competition.
 Similarly, it could make sense for the U.S.
-to be focusing on improving IF in countries with already low levels, as its aid could be more effective here...
-This opposing logic generates the following hypotheses:
+to be focusing on improving the internet in countries with relatively illiberal models, as it is plausible that foreign aid's influence on the internet would have higher marginal benefits.
+This opposing set of logic generates the following hypotheses:
 \begin{align*}
     &\text{``Intrastate Competition Hypothesis''}\\
-    H_{1_b}&:\;\text{U.S. aid directly related to Chinese aid}\\
+    H_{1b}&:\;\text{U.S. aid directly related to Chinese aid}\\
     &\text{``Bottom Up Hypothesis''}\\
-    H_{2_b}&:\;\text{U.S. aid inversely related to IF}\\
+    H_{2b}&:\;\text{U.S. aid inversely related to recipient internet model}\\
 \end{align*}
@@
 I chose to focus on Africa during this time period for three main reasons.
-Firstly, Africa has been a major area of focus for China finance projects, with the Council on Foreign Relations claiming that ``China already provides more financing for information and communications technology than all multilateral agencies and leading democracies combined do across the continent'' (CFR).
-Secondly, data for Chinese spending ends in 2017.
-Thirdly, while 2010 is indeed an arbitrary starting point, both Chinese aid and the development of the internet diminish greatly before this period.
+Firstly, Africa has been a major area of focus for China's finance projects, with the Council on Foreign Relations claiming that ``China already provides more financing for information and communications technology than all multilateral agencies and leading democracies combined do across the continent'' (\cite{kurlantzicketal.2020}).
+Secondly, the data I use for Chinese spending ends in 2017.
+Thirdly, while 2010 is indeed an arbitrary starting point, it seems like a reasonable choice since both Chinese aid and the internet diminish greatly before this period.
 
@@
 
-The DSP is one of two indices measuring a political concept of internet usage and data from the DSP is potentially superior for measuring the concept of internet freedom compared to the primary alternative, Freedom on the Net (FOTN, \cite{house2022}).
-FOTN features frequent missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.
+The DSP is one of two indices measuring a political concept of internet usage and data from the DSP is potentially superior for measuring the concept of internet liberalism in Africa compared to the primary alternative, Freedom on the Net (FOTN, \cite{freedomhouse2022}).
+FOTN frequently features missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.
 
@@
 This is the one of the only datasets on Chinese aid and is certainly the largest and most specific.
-Aid is quantified as official financial and in-kind commitments and is measured in constant 2017 USD.
+Aid is quantified as official financial and in-kind commitments and is collected using a methodology designed to track under-reported aid flows, such as those from China.
 
 %U.S. aid
-U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{government2022e}). It is measured in constant 2022 USD.
+U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{u.s.agencyforinternationaldevelopment2022}). This data was more up-to-date than alternatives.
 
 %controls
-I include gross domestic product (GDP) per capita, an indicator for internet development, and political freedom as controls.
+For controls, I include gross domestic product (GDP) per capita, an indicator for internet development, and a measure for political freedom as controls.
 Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.
 
-GDP per capita is assessed in constant 2021 USD and as is supplied by the World Bank (\cite{bank2022}).
-The indicator for internet development, internet usage by country, is calculated as a percentage and is provided by the International Telecommunication Union (\cite{itu2022}).
-Lastly, the chosen measure for political freedom comes from V-Dem (\cite{coppedge2022a}).
+GDP per capita is supplied by the United Nations (\cite{unitednationsstatisticsdivision2019}) while the indicator for internet development is provided by the International Telecommunication Union (ITU, \cite{itu2022}).
+This indicator is a proxy for how developed the internet is in a country and represents the percentage of the population using the internet.
+Lastly, the chosen measure for political freedom comes from Polity (\cite{marshall2018}).
+
+Using a GDP deflator (\cite{organizationforeconomicco-operationanddevelopment2022}), the values for U.S. aid, Chinese aid, and GDP per capita are all in constant 2015 USD.
 
 %Missing data
-Of the 10 African countries I am missing, the DSP is missing values for Somalia, Eritrea, Sudan, and South Sudan, which might make sense given that Somalia and South Sudan were both at civil war during the observation period while Sudan and Eritrea have experienced instability after previous wars.
+Of the 10 African countries I am missing data for, the DSP is missing values for Somalia, Eritrea, Sudan, and South Sudan, which might make sense given that Somalia and South Sudan were both at civil war during the observation period while Sudan and Eritrea have experienced instability after previous wars.
 Similarly, Libya was missing data from the ITU following the Western invasion of Libya in 2011.
@@
 Descriptive statistics can be seen in Table 1.
+As might be able to be seen here, logged Chinese aid is a quirky variable.
+It is the only variable subject to influential cases, since it contains many zeroes but several large values.
+I am not sure how else to treat this variable other than having logged it to minimize the scale of the distance between the values.
+All other variables contain no statistically significant influential cases and seem relatively well behaved.
 
@@
 A correlation matrix of the variables included in my regressions is given in Figure 2.
-As can be seen, there are a couple relatively strong relationships between variables.
-The variable for regime type, ``Polity 2,'' has a correlation of 0.62 with ``Internet Freedom'' while ``Logged GDP per Capita'' has correlation of 0.65 with the variable measuring internet development, ``Percent Using Internet.'' While possible, I do not think that these correlations are high enough to warrant concerns over multicollinearity or the lack of linear independency in the data.
-Interestingly, ``Logged U.S.
-Aid'' appears somewhat negatively correlated with ``Logged GDP per Capita'' and ``Percent Using Internet'' is somewhat positively correlated to ``Internet Freedom.''
+As can be seen, there are a couple relatively strong relationships between the variables.
+The variable for regime type, Polity, has a correlation of 0.62 with the internet freedom variable while logged GDP per capita has a correlation of 0.65 with the variable representing the percent of the population using the internet.
+While possible, I do not think that these correlations are high enough to warrant concerns over multicollinearity or the lack of linear independency in the data.
+Interestingly, logged U.S.
+Aid appears somewhat negatively correlated with logged GDP per capita and variable for the percent using the internet is somewhat positively correlated to that for internet freedom.
+These correlations are relatively small, but it could be interesting to explore the relationship between U.S.
+aid and GDP.
 
@@
 Table 2 displays regression results from a model with and without an interaction between internet freedom and logged Chinese foreign aid.
-
 \subsection*{Table 2}
@@
 \end{table}
+\pagebreak
+
+The only covariates that appear statistically significant are the two main explanatory variables.
+Both internet freedom and logged Chinese aid seem to increase logged U.S.
+aid on average when accounting for the other covariates.
+To gauge the substantive size of the coefficients, I plot the marginal effect of internet freedom on logged U.S.
+aid in Figure 3 and the marginal effect of logged Chinese aid on logged U.S.
+aid in Figure 4.
+
+Since the effect of internet freedom appears directly related to logged U.S.
+aid, there is more empirical support for $H_{2a}$, the ``Trickle Down Hypothesis,'' than $H_{2b}$, the ``Bottom up Hypothesis.'' Support for $H_{2a}$ implies that I cannot reject the notion that the U.S.
+is allocating foreign aid to countries with higher levels of internet freedom.
+
+The ``Intrastate Competition Hypothesis,'' or $H_{1b}$, finds more support than the ``Interstate Competition Hypothesis,'' $H_{1a}$, since logged Chinese aid is directly related to logged U.S.
+aid.
+It seems that $H_{1b}$ cannot be rejected, and that the U.S.
+is plausibly spending more foreign aid in countries where China is also spending more.
+
+Interestingly, there is a lack of support for an interaction between internet freedom and logged Chinese aid.
+Apparently, the effects of a recipient's internet freedom level and the amount of Chinese aid they are receiving have completely separate effects on U.S.
+aid.
+However, it is unclear how consequential this lack of support for the interaction term is, since both variables by themselves are significant and two-way fixed effects are used.
 
 \pagebreak
-Seems to support for $H_{1_b}$ and $H_{2a}$.
-In other words, seems that the U.S., in response to Chinese foreign aid, is using a intrastate competition strategy in competing with China and trying to increase IF in a trickle down manner.
-
-\pagebreak
-Figure 3 shows the marginal effect of internet freedom on logged U.S. foreign aid.
+Figure 3 shows the marginal effect of internet freedom on logged U.S.
+foreign aid.
+The coefficient on the internet freedom variable is 0.0206, implying that a one unit increase in a recipient's internet freedom level is associated with an increase in U.S.
+aid by 2.08 percent on average, accounting for other covariates.
+This appears to be a substantively large effect.
 
@@
 \pagebreak
-Figure 4 graphs the marginal effect of logged Chinese foreign aid on logged U.S. foreign aid.
+Figure 4 graphs the marginal effect of logged Chinese foreign aid on logged U.S.
+foreign aid.
+The coefficient on the variable for logged Chinese foreign aid is 0.0058, implying that a one percent increase in Chinese aid is associated with a 0.0058 percent increase in U.S.
+aid on average, accounting for other covariates.
+This does not appear to be a substantively large effect.
 
@@
 \section*{Discussion}
-Talk about implications of trickle-down, intrastate strategy...
+The support found for both $H_{1b}$ and $H_{2a}$ is a little surprising for the following reason.
+While it appears that the direct relationship between internet freedom and U.S.
+aid is in accordance with the literature ($H_{2a}$), this relationship was hypothetically built on the previous finding that U.S.
+aid is relatively ineffective in countries receiving Chinese aid (\cite{dreher2021}).
 
-Why U.S. might be interested in IF.
+Subsequently, it should also be the case that U.S.
+aid is inversely related to Chinese aid ($H_{1a}$) if the U.S.
+rationally desires its aid to be effective, which should mean that aid would be going to countries with more liberal internet systems due to the absence of illiberal effects from Chinese aid.
+However, U.S.
+aid seems to also be directly related to Chinese aid, undermining this logic.
+Thus, it seems that the U.S.
+either does not highly value the effectiveness of its aid or that its aid is more effective at improving internet liberalism where China is also spending money.
 
-Is U.S. response effective at keeping China out? Probably not, although they are ramping up efforts.
+Regardless, it is interesting that some support was found for both the ``Trickle Down Hypothesis'' and the ``Intrastate Competition Hypothesis.'' Given the lack of an interaction term, it would seem that the U.S.
+is reacting to the influx of Chinese aid by increasing foreign aid spending in countries where either Chinese aid is also present or the recipient has a relatively liberal internet model, but not both.
+Further investigation of this phenomenon in the data is required.
 
 \section*{Conclusion}
+In this article I argue that it is rational for the U.S.
+to respond to a global increase in Chinese foreign aid by changing its own aid allocation strategy and focusing on how recipients treat the internet.
+In particular, the literature seems to imply that the U.S.
+should concentrate spending in countries devoid of Chinese aid and with highly liberal models of internet governance.
+
+Analyzing 44 African countries from 2010-2017, I find mixed results for this set of expectations.
+While U.S.
+aid indeed seems positively related to a measure of internet governance, it is also directly related to Chinese foreign aid.
+
+Given these findings, I suggest two future lines of research on this topic.
+Firstly, more rigorous research needs to be conducted on the true relationship between China's aid projects and the effects on internet governance.
+Much of the existing discussion on this area is non-scientific and biased towards U.S.
+interests.
+The results shown here do not directly speak to how Chinese foreign aid affects recipient internet usage, but there was a lack of support for the hypothesis that U.S.
+aid and Chinese aid should be inversely related ($H_{1a}$).
+Perhaps these countries do not have entirely opposite, neatly ideological goals concerning how countries use the internet around the world.
+
+Secondly, depending on the strength of these results, there should be further work done explicating why exactly the U.S.
+is interested in boosting internet liberalism in Africa.
+If the U.S.
+is truly interested in increasing democracy and liberalism around the world in general, it is unclear if this motivation is sufficient to account for such large expenditures if the U.S.
+is a rational actor.
+In other words, to the extent that the internet is relevant for the foreign aid decision of the U.S., it is worthwhile pursuing explanations for what the U.S.
+materially gains from this practice.
 
@@
 \section*{Appendix}
-The entry game described in the theory section (Figure 1) is solved via backwards induction. If $China$ chooses to $Enter$, the $Donor$ chooses to $Spend$ if and only if (iff):
+The entry game described in the theory section (Figure 1) is solved via backwards induction.
+The parameter $\delta$ inversely represents the influence that U.S.
+aid has on a recipient's internet model while $\gamma$ inversely represents the influence of Chinese aid.
+In other words, low values of $\delta$ or $\gamma$ imply high degrees of influence.
+
+Hypothetically, if $China$ chooses to $Spend$, the $U.S.$ chooses to $Spend$ if and only if (iff):
 
 \begin{tabular}{c l}
-    $\frac{\pi}{\gamma}-c > -1$ &  Condition for $Donor$ to $Spend$, from the payoffs on $Spend$ and $\neg Spend$, respectively.\\
-    $\frac{\pi}{c-1} > \gamma $ & Isolating $\gamma$.\\
+    $\frac{\gamma}{\delta}-c_2> \delta$ &  As $\delta$ increases, this condition becomes harder to satisfy.\\
 \end{tabular}
 
-Where $\gamma \in (0,1]$. As $\gamma$ gets arbitrarily small, $\pi$ goes to infinity.
+Hypothetically, if $China$ chooses to $\neg Spend$, the $U.S.$ chooses to $Spend$ iff:
+
+\begin{tabular}{c l}
+    $\frac{1}{\delta}-c_2> \delta$ &  As $\delta$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Hypothetically, when $\delta$ is arbitrarily small, the $U.S.$ chooses to $Spend$ and $China$ $Spend$s iff:
+
+\begin{tabular}{c l}
+    $\frac{\delta}{\gamma}-c_1> \gamma$ &  As $\gamma$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Hypothetically, when $\delta$ is arbitrarily large, the $U.S.$ chooses to $\neg Spend$ and $China$ $Spend$s iff:
+
+\begin{tabular}{c l}
+    $\frac{1}{\gamma}-c_1> \gamma$ &  As $\gamma$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Summarily, if $\delta$ and $\gamma$ are arbitrarily small, both players will $Spend$.
+When $\delta$ is arbitrarily small and $\gamma$ is arbitrarily large, the $U.S.$ $Spend$s and $China$ $\neg Spend$.
+If $\delta$ is arbitrarily large and $\gamma$ is arbitrarily small, the $U.S.$ $\neg Spend$ and $China$ $Spend$s.
+When $\delta$ is arbitrarily large and $\gamma$ is arbitrarily large, neither player $Spend$s.
 

### 2022-12-06 Extra.tex
rough draft
draft

@@
-%----------------------------------------------------------
-%To do
-Case study comparing U.S.
-aid changes to countries that had similar internet freedom levels but experienced different influxes of Chinese aid?
-Need to do case studies of results and examples, in general.
-%----------------------------------------------------------
-%Rough Draft extras
-\section*{Literature Review}
+\section{Final Draft Extras}
+\subsection{Introduction}
+The People's Republic of China (China) has become the ``lender of first resort'' for developing countries in the last decade \parencite[1]{dreher2022}.\footnote{All data and code can be found on GitHub at https://github.com/nnray/Analysis.
+In the ``Code'' folder, navigate to ``628.Rmd'' for my R script.
+My full \LaTeX \;file can be found at https://github.com/nnray/628 under ``Paper.tex''.} An expanse of recent work has begun documenting how our understanding of global finance has changed since China's financial preeminence.
+Some of this research looks at how Western money, both foreign aid and loans, has been affected by the unprecedented availability of Chinese money (e.g., \cite{hernandez2017}).
+With regard to the United States (U.S.), there is evidence that its foreign aid has become more effective at stimulating liberal democratic values among recipient citizens (\cite{blair2022}).
+%other work
+
+Surprisingly, though, there has been relatively little work on how U.S.
+spending behavior itself may have changed with the rise of Chinese finance.
+In this article I aim to address this shortcoming and explore how U.S.
+foreign aid spending has responded to a global increase in Chinese finance.
+Focusing on Africa, I expect that an empirical relationship between U.S.
+aid and internet governance has emerged as a result of Chinese spending.
+Using the most rigorous measure of internet governance available, I find evidence that internet governance describes the variation in U.S.
+foreign aid spending after accounting for common measures of regime type, economic development, and internet development.
+
+I make two contributions in the uncovering of this relationship.
+Firstly, I add to the foreign aid literature by detailing a novel determinant of U.S.
+foreign aid and providing another clue on how Chinese money may have impacted Western donor behavior more generally.
+Secondly, I speak to literature on the liberal international order (LIO) and the concern that China may undermine tenets of the LIO.
+%expand on this a bit?
+
+I proceed as follows.
+In the next section, I define internet governance and detail the theoretical reasoning behind its importance and relationship to U.S.
+foreign aid.
+After developing hypotheses, I describe my empirical strategy and data.
+My findings and a discussion of their implications are presented before concluding.
+
+\subsection{Theory}
+Internet governance was defined in the second phase of the World Summit on the Information Society (WSIS), a meeting led by the United Nations' (UN) International Telecommunication Union (ITU), which was held in Tunis, Tunisia in 2005.
+It is ``the development and application by governments, the private sector and civil society, in their respective roles, of shared principles, norms, rules, decision-making procedures, and programmes that shape the evolution and use of the Internet'' (World Summit on the Information Society 2005, item 34).
+Essentially, internet governance refers to the way in which governments, firms, and citizens interact with the internet.
+%What it composes of and commmon measures...
+
+But why focus on internet governance?
+What can be uniquely gained in looking at U.S.
+foreign aid through a potentially new determinant?
+The U.S.
+and China have recently been observed to direct a substantial amount of money towards building digital infrastructure in the developing world, which consists of the cables and other hardware that facilitate internet connectivity.
+For example, (\cite{thewhitehouse2022}) and (\cite{shen2022}).
+
+I contend that these observations are difficult to fully explain by previous accounts.
+Past work has identified that foreign aid is dominated by the self-interests of donors and is not allocated according to recipient need (\cite{hoeffler2011}; others).
+That is, foreign aid expenditures on digital infrastructure is likely in the foreign policy interests of the U.S..
+However, it is not ex ante obvious how exactly this is the case.
+The U.S.
+has historically prioritized economic growth (\cite{dreher2021}; others), but a desire to stimulate economic growth does not explain the adversarial nature of this recent digital infrastructure aid and the apparently zero-sum stance against China.
+Indeed it can be argued that Chinese investment is in the U.S.'s interests if those interests are purely concerned with economic growth.
+Political competition likely is involved here, but it has yet to be shown how investing in digital infrastructure is related to the political competition between the U.S.
+and China, although it is possible that the U.S.
+has returned to a Cold-War style of foreign policy.
+Also, the U.S.
+has previously proven that they are perfectly willing to aid those with illiberal political models (Saudi Arabia?).
+Consequently, I speculate that an adequate explanation of the recent competition regarding digital infrastructure likely revolves around the digital aspect itself, and the U.S.
+China competition over technological superiority.
+This is manifested in concern for internet governance.
+Thus, I argue that U.S.
+foreign aid should be empirically related to internet governance for the following reasons.
+
+Summarily, it is rational for the U.S.
+to condition its foreign aid spending on a recipient's internet governance in response to a rise in Chinese money because: 1) the U.S.
+has increasingly voiced a preference for more liberal internet governance worldwide; 2) U.S.
+foreign aid has been shown to be effective at promoting liberal values; 3) the Chinese are suspected of sharing technology with recipient countries and negatively influencing internet governance; and 4) the U.S.
+and China have been engaged in a competition for technological superiority, of which the internet is a part of.
+To the extent that these statements are true, together they imply that it makes sense for the U.S.
+to spend aid in a way that positively influences the liberalness of internet governance and inhibits the spread of Chinese technology and influence.
+I will detail the support behind these statements in the following paragraphs.
+
+Firstly, there is qualitative evidence that the U.S.
+has a preference for liberal internet governance in other countries.
+In 1997 President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (The White House 1997, principles 1-5) and by the 2000's, these preferences crystallized into foreign policy initiatives as so-called ``internet freedom'' programs were initiated by the U.S.
+Department of State (e.g., \cite{henry2014}; \cite{u.s.departmentofstate2021}).
+U.S.
+attention to internet governance has continued to grow, with Secretary of State Hillary Clinton outlining U.S.
+demands for a ``free'' internet (\cite{u.s.departmentofstate2010}) and the U.S.
+Senate considering additional funding for internet freedom programs in 2022 (\cite{foreignrelationscommittee2022}).
+In the same year, a declaration led by the U.S.
+was signed by over 60 countries to combat global ``digital authoritarianism'' (\cite{u.s.departmentofstate2022}).
+
+Secondly, there is indirect evidence that foreign aid is an effective tool for influencing internet politics in recipient countries.
+For instance, in Africa U.S.
+aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents when compared to Chinese foreign aid \cite{blair2022}.
+%bearce and tirone, dreher et al, etc
+
+Thirdly, there is reason to believe that Chinese aid could influence a recipient's political use of the internet in an illiberal manner.
+China launched the Belt and Road Initiative (BRI) in 2013 and the Digital Silk Road (DSR) in 2015, both aimed at enlarging China's global influence while strengthening economic growth (\cite{dreher2022}).
+In doing so, it is alleged that China is sharing technology with recipient states that could improve state capacity and allow recipients to curb free expression, both digitally and physically (\cite{hillman2021}).
+Other work, though, rightly points out that China is a non-unitary actor with private firms that could undermine a state-led technology sharing campaign (\cite{shen2018a}) and that China's technology spending has not changed much immediately following the announcement of the DSR (\cite{tugendhat2021}).
+The West believes China to be spreading its illiberal model of the internet (e.g., \cite{hillman2021}; \cite{u.s.departmentofstate2022}), or at least that there is a good chance for it to (\cite{triolo2020}).
+
+Fourthly, it has been observed by some experts that the U.S. and China are competing for global technological superiority (\cite{hass2021}).
+
+Convincingly explain how these four statements lead one to expect that U.S.
+aid is empirically related to internet governance.
+Describe diagram 1.
+Then state this expectation formally, think about directionality of hypotheses and how to base them more off the literature (effectiveness, etc.).
+
+\pagebreak
+%may add Time 0 about Soviet influence to emphasize the similarity here in U.S. response to China?
+\subsection*{Diagram 1}
+    %Time 0
+    \begingroup
+    \leftskip=0cm plus 0.5fil \rightskip=0cm plus -0.5fil
+    \parfillskip=0cm plus 1fil
+    Time 1 (1990's - 2000's)\par
+    \endgroup
+    \begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+    [node distance = 1cm, auto,font=\footnotesize,
+    % STYLES
+    every node/.style={node distance=3.5cm},
+    % The comment style is used to describe the characteristics of each force
+    comment/.style={rectangle, inner sep= 5pt, text width=4cm, node distance=0.25cm, font=\scriptsize\sffamily},
+    % The force style is used to draw the forces' name
+    force/.style={rectangle, draw, fill=black!10, inner sep=5pt, text width=4cm, text badly centered, minimum height=1.2cm, font=\bfseries\footnotesize\sffamily}]
+    % Draw forces
+    \node [force] (usaid) {U.S. Aid};
+    \node [force, above of=usaid] (effectiveness) {Effectiveness};
+    \node [force, left=1cm of effectiveness] (preferences) {Preference for Politico-economic Liberalism};
+    %preferences
+    \node [comment, below=0.25cm of preferences] {Preference for a liberal internet largely subsumed under politico-economic liberalism and not independently influential};
+    %effectiveness
+    \node [comment, below=0.2cm of effectiveness] {Aid positively effective on economic growth (Bearce and Tirone 2010)};
+    %usaid
+    \node [comment, below=0.25cm of usaid] {Aid primarily directed to recipients with more liberal politico-economic structures};
+    % Draw the links between forces
+    \path[->,thick] 
+    (effectiveness) edge (usaid)
+    (preferences) edge (usaid);
+    \end{tikzpicture} 
+    \end{figure}
+
+    %Time 1
+    \begingroup
+    \leftskip=0cm plus 0.5fil \rightskip=0cm plus -0.5fil
+    \parfillskip=0cm plus 1fil
+    Time 2 (2010's)\par
+    \endgroup
+    \begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+    [node distance = 1cm, auto,font=\footnotesize,
+    % STYLES
+    every node/.style={node distance=3.5cm},
+    % The comment style is used to describe the characteristics of each force
+    comment/.style={rectangle, inner sep= 5pt, text width=4cm, node distance=0.25cm, font=\scriptsize\sffamily},
+    % The force style is used to draw the forces' name
+    force/.style={rectangle, draw, fill=black!10, inner sep=5pt, text width=4cm, text badly centered, minimum height=1.2cm, font=\bfseries\footnotesize\sffamily}]
+    % Draw forces
+    \node [force] (usaid) {U.S. Aid};
+    \node [force, above of=usaid] (effectiveness) {Effectiveness};
+    \node [force, left=1cm of effectiveness] (preferences) {Preference for Politico-economic Liberalism \underline{and} a Liberal Internet};
+    \node [force, above of=preferences] (caid) {Chinese Aid};
+    %preferences
+    \node [comment, below=0.25cm of preferences] {Preference for a liberal internet largely now independently influential due to technology competition with China};
+    %effectiveness
+    \node [comment, below=0.25cm of effectiveness] {Aid now positively effective at promoting liberal values in light of Chinese aid};
+    %usaid
+    \node [comment, below=0.25cm of usaid] {Aid now related to a recipient's internet liberalness};
+    %caid
+    \node [comment, below=0.25cm of caid] {Chinese launched the BRI in 2013 but had been increasing its foreign aid before then};
+    % Draw the links between forces
+    \path[->,thick] 
+    (effectiveness) edge (usaid)
+    (preferences) edge (usaid)
+    (caid) edge (preferences)
+    (caid) edge (effectiveness);
+    \end{tikzpicture}
+    \end{figure}
+
+    %Time 2
+\pagebreak
+
+\section{Rough Draft Extras}
+\subsection*{Literature Review}
 Foreign aid before BRI (a few sentences with biggest takeaways; self-interest, ineffectiveness)
@@
 
-\section*{Theory}
+\subsection*{Theory}
 To explain how other donor countries should be expected to react to China's recent dominance of the international financing arena, I model a simple entry game between China and an unspecified country.
@@
 
-\section*{Conclusion}
+\subsection*{Model}
+I argue that it is rational for the U.S.
+to consider the internet when distributing foreign aid and, specifically, that a recipient's current use of the internet should be a strong predictor for receiving U.S.
+funds.
+To explore the rationality of this alleged behavior, I consider a simple entry deterrence game.
+China is the first-mover and decides to enter or spend foreign aid in a particular country by allocating aid or not.
+The U.S.
+then moves second, making a similar decision to distribute foreign aid depending on if China has entered or not.
+There are two parameters that decide the payoffs of the game: $\gamma$, the influence that Chinese aid has on a recipient country's internet model and $\delta$, the influence of U.S.
+aid.
+Low values represent high degrees of influence for both.
+Essentially, the strategies of the players are symmetrical, with each country deciding to spend aid if their degree of influence is relatively larger than their rival's.
+
+When U.S.
+foreign aid is net effective at encouraging a liberal model of the internet in a country, it decides to allocate aid.
+China, observing the U.S.'s expenditure, decides not to distribute foreign aid in that country.
+However, if the U.S.
+concludes not to spend aid monies due to its aid being relatively ineffective at influencing a recipient's use of the internet, then China allocates aid and exerts its influence on a recipient's internet usage.
+The extensive-form of this game is displayed in Figure 1, with more detail given in the appendix.
+
+\subsubsection*{Figure 1}
+\begin{istgame}[scale=4]
+\xtdistance{5mm}{15mm}
+\istroot(0)(0,0){\textit{China}}
+\istb{\neg Spend}[al]
+\istb{Spend}[ar]{0\leq\gamma,c,\delta\leq1}[below,yshift=25mm,xshift=25mm]
+\endist
+\xtdistance{5mm}{10mm}
+\istroot(1)(0-1)<above left>{\textit{U.S.}}
+\istb{\neg Spend}[l]{(\gamma,\;\delta)}
+\istb{Spend}[r]{(\gamma,\;\frac{1}{\delta}-c_2)}
+\endist
+\istroot(2)(0-2)<above right>{\textit{U.S.}}
+\istb{\neg Spend}[l]{(\frac{1}{\gamma}-c_1,\;\delta)}
+\istb{Spend}[r]{(\frac{\delta}{\gamma}-c_1,\;\frac{\gamma}{\delta}-c_2)}
+\endist
+\xtdistance{10mm}{10mm}
+\end{istgame}
+
+\subsection*{Notable Assumptions}
+Ample explanation is in order.
+Firstly, there is a host of technical restrictions inherent to my argument.
+Other determinants of foreign aid, for example, are not considered here.
+Thus, any changes in spending should be interpreted as the U.S.
+or China deciding to spend on internet influence or not $ceteris\;paribus$, or holding other foreign aid avenues constant.
+
+These decisions are also stipulated as dichotomous, where China or the U.S.
+simply agree to spend foreign aid or not.
+In reality, foreign aid spending is clearly continuous in nature and, moreover, typically based on repeated interactions between both other donors and recipients.
+This is also a full information model, with no uncertainty present over players' decisions to spend.
+I further assume that payoffs from the decision to spend or not are zero-sum at the extremes, where an arbitrarily large influence by one country dwarfs the incentives for the other to spend foreign aid.
+
+Secondly, and most critically, there are a number of substantive assumptions invoked in this theoretical exposition that must be detailed.
+For one, I am assuming that the U.S.
+has a preference over internet governance around the world and, additionally, that foreign aid is an effective instrument for the U.S.
+to actualize this preference.
+Somewhat symmetrically, I assume that Chinese aid has an effect on a recipient's internet model and that this effect is negative or illiberal.
+I will take these assumptions in turn, connecting these behavioral claims to the established literature where possible.
+
+\subsection*{Implications}
+Returning to my theory, the model simplistically states that the U.S.
+and China should allocate foreign aid when aid is relatively effective at influencing a recipient's political model of the internet.
+Some of the literature implies that U.S.
+aid is effective when Chinese aid is low or not present (\cite{dreher2021}).
+Since it is assumed that Chinese aid has a negative effect on how recipients use the internet, this further implies that U.S.
+aid should be spent in countries with more liberal models of the internet to ensure maximum effectiveness.
+These concurrent implications lead to the following hypotheses:
+
+\begin{align*}
+    &\text{``Interstate Competition Hypothesis''}\\
+    H_{1a}&:\;\text{U.S. aid inversely related to Chinese aid}\\
+    &\text{``Trickle Down Hypothesis''}\\
+    H_{2a}&:\;\text{U.S. aid directly related to recipient internet model}\\
+\end{align*}
+
+I label $H_{1a}$ the ``Interstate Competition Hypothesis'' because if U.S.
+foreign aid is truly most effective where Chinese aid is not present, this implies that the U.S.
+would respond to an increase in Chinese aid by focusing its spending in countries not currently receiving aid from China.
+China and the U.S.
+would thus be spending in completely separate countries or regions.
+
+$H_{2a}$ is titled the ``Trickle Down Hypothesis'' since the U.S.
+would spend more foreign aid in countries with more liberal internet models, implying that the U.S.'s preference for a globally liberal internet model is manifested by it focusing on bolstering countries with already relatively liberal models.
+
+However, it is also hypothetically possible that the U.S.
+would try to compete with Chinese aid by directly countering it in countries where the Chinese are active and generating intrastate competition.
+Similarly, it could make sense for the U.S.
+to be focusing on improving the internet in countries with relatively illiberal models, as it is plausible that foreign aid's influence on the internet would have higher marginal benefits.
+This opposing set of logic generates the following hypotheses:
+\begin{align*}
+    &\text{``Intrastate Competition Hypothesis''}\\
+    H_{1b}&:\;\text{U.S. aid directly related to Chinese aid}\\
+    &\text{``Bottom Up Hypothesis''}\\
+    H_{2b}&:\;\text{U.S. aid inversely related to recipient internet model}\\
+\end{align*}
+
+\subsection*{Discussion}
+test if Chinese aid related to IF itself, implications
+
+\subsection*{Conclusion}
 Crazy conclusion: I suspect that the U.S.
@@
 
-%----------------------------------------------------------
-%Old theory ideas (explaining an expected relationship)
-\section*{Theory} 
+\subsection*{Appendix}
+The entry game described in the theory section (Figure 1) is solved via backwards induction.
+The parameter $\delta$ inversely represents the influence that U.S.
+aid has on a recipient's internet model while $\gamma$ inversely represents the influence of Chinese aid.
+In other words, low values of $\delta$ or $\gamma$ imply high degrees of influence.
+
+Hypothetically, if $China$ chooses to $Spend$, the $U.S.$ chooses to $Spend$ if and only if (iff):
+
+\begin{tabular}{c l}
+    $\frac{\gamma}{\delta}-c_2> \delta$ &  As $\delta$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Hypothetically, if $China$ chooses to $\neg Spend$, the $U.S.$ chooses to $Spend$ iff:
+
+\begin{tabular}{c l}
+    $\frac{1}{\delta}-c_2> \delta$ &  As $\delta$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Hypothetically, when $\delta$ is arbitrarily small, the $U.S.$ chooses to $Spend$ and $China$ $Spend$s iff:
+
+\begin{tabular}{c l}
+    $\frac{\delta}{\gamma}-c_1> \gamma$ &  As $\gamma$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Hypothetically, when $\delta$ is arbitrarily large, the $U.S.$ chooses to $\neg Spend$ and $China$ $Spend$s iff:
+
+\begin{tabular}{c l}
+    $\frac{1}{\gamma}-c_1> \gamma$ &  As $\gamma$ increases, this condition becomes harder to satisfy.\\
+\end{tabular}
+
+Summarily, if $\delta$ and $\gamma$ are arbitrarily small, both players will $Spend$.
+When $\delta$ is arbitrarily small and $\gamma$ is arbitrarily large, the $U.S.$ $Spend$s and $China$ $\neg Spend$.
+If $\delta$ is arbitrarily large and $\gamma$ is arbitrarily small, the $U.S.$ $\neg Spend$ and $China$ $Spend$s.
+When $\delta$ is arbitrarily large and $\gamma$ is arbitrarily large, neither player $Spend$s.
+
+\section{Other Extras}
+\subsection*{Theory} 
 U.S. wants to protect American Big Tech markets that are potentially going to be lost in a zero-sum game with Chinese companies (if China successfully shares tech and builds infrastructure).
@@
 
-\section*{Alternative Explanations}
+\subsection*{Alternative Explanations}
 1) The U.S. is not concerned necessarily about Big Tech markets, but instead increasing aid in an attempt to thwart China's goals of economic growth abroad.
@@
 9/29
-\section*{Theory}
+\subsection*{Theory}
 Depending on how the literature review goes, we can mention our theory of why it is rational for the U.S.
 to engage in promoting internet freedom (aid effectually increases democracy, and democracy is good for business (American Big Tech)).
-
-%----------------------------------------------------------

### 2022-12-06 Paper.tex
draft

--- 2022-09-30 Paper.tex
@@
-\documentclass{article}
+\documentclass[12pt]{article}
 \usepackage[utf8]{inputenc}
-\title{\vspace{-2cm}Digital Authoritarianism and United States Foreign Aid\vspace{-0.5cm}}
+\title{\vspace{-2.75cm}Conflating Global Finance with Technological Competition: The U.S.'s Reaction to Chinese Foreign Aid\vspace{-0.5cm}}
 \author{Nicholas Ray}
-\date{\vspace{-0.5cm}September 2 2022\vspace{-1cm}}
+\date{\vspace{-0.30cm}December 7 2022\vspace{-1cm}}
 \usepackage[margin=1in]{geometry}
@@
 \doublespacing
+\usepackage{footmisc}
+\renewcommand{\footnotelayout}{\setstretch{1.75}}
+\usepackage{footmisc}
+\renewcommand{\footnotesize}{\normalsize}
+\usepackage{caption}
+\usepackage{tikz}
+\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains}
+\usepackage{blindtext}
+\usepackage{istgame}
 \usepackage[backend=biber, style=authoryear, maxbibnames=99,uniquelist=false]{biblatex}
@@
 \maketitle
+\begin{abstract}
+    Emphasize policy! \blindtext
+\end{abstract}
+
+\textit{Keywords}: Foreign Aid, Internet Governance, Africa
+
 \section*{Introduction}
-The United States (U.S.) has been openly concerned about the internet since at least 1997, when President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (\cite{government1997}).
-By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as internet freedom programs cropped up in different facets of the U.S.
-government, such as the Department of State (e.g., \cite{government2021}) and the U.S.
-Agency for Global Media (\cite{government2022d}).
-U.S.
-attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).
+\footnote{All data and code can be found on GitHub at https://github.com/nnray/Analysis.
+In the ``Code'' folder, navigate to ``628.Rmd'' for my R script.
+My full \LaTeX \;file can be found at https://github.com/nnray/628 under ``Paper.tex''.}
 
-Today internet freedom has seemingly reached a crescendo in U.S.
-foreign policy.
-Bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S.
-and over 60 other countries to combat ``digital authoritarianism'' (\cite{government2022b}), and an unprecedented, U.S.-led G7 spending plan was unveiled to improve physical and digital infrastructure abroad (\cite{government2022c})- all in 2022.
+\section*{What Needs Explaining?}
+%U.S. digital infrastructure spending, other aid related to ICT/internet
+The U.S.
+has been increasingly funding digital infrastructure in countries around the world.
+In 2022, for example, the U.S.
+initiated the Partnership for Global Infrastructure and Investment (PGII) in collaboration with the other members of the Group of Seven (G7) that aims to spend \$600 billion on global infrastructure projects over the next five years (\cite{thewhitehouse2022}).
+Included in the PGII is a \$600 million contract to a U.S.
+telecommunications company to build submarine internet cables that connect Singapore to France via Egypt and the Horn of Africa.
+%other U.S.
+examples?
+IF programs?
 
-Along the way, the People's Republic of China (China) and the Russian Federation (Russia) have been named the antagonists to global internet freedom and the explicit motivations behind the U.S.'s focus on the subject.
-During this intensifying period of U.S.
-apprehension of digital authoritarianism, China launched it's Belt and Road Initiative (BRI) in 2013 and the accompanying Digital Silk Road Initiative (DSR) in 2015.
-Collectively, these projects have catalyzed over one trillion U.S.
-dollars in developing-country investment with the intent to improve China's perception and influence in the world while fostering it's domestic economic growth (\cite{dreher2022}).
+%Chinese digital infrastructure spending
+This intention of the U.S.
+to build digital infrastructure closely resembles the behavior of China over the past two decades.
+Since engaging in their ``going out'' strategy of investing overseas to promote domestic economic growth, China has turned to funding massive infrastructure projects through its Belt and Road Initiative (BRI) and becoming the ``lender of first resort'' for developing countries (\parencite[1]{dreher2022}).
+Two years after announcing the BRI, China launched the largely symbolic (\cite{tugendhat2021}) Digital Silk Road (DSR) that still encapsulates China's desire to build global digital infrastructure favorable to them (\cite{shen2022}).
+Some examples of relevant Chinese projects are shown in table 1 \nocite{agbebi2022}.
 
-Ample academic work has recently attempted to outline the effects of China's aid program (e.g., \cite{gehring2022}; \cite{dreher2018}; \cite{blair2021}).
-Yet, little research has explored the potential effects of China's spending on U.S.
-foreign policy.
-In this paper, I seek to understand if U.S.
-foreign aid, an important aspect of foreign policy, can be understood as a response to China's aid ambitions.
-Specifically, I hypothesize that U.S.
-foreign aid to Africa is related to a recipient's internet freedom level and their influx of Chinese aid.
+\subsubsection*{Table 1}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.449]{Figures/ChineseAid.png}
+    \captionof{table}{Agbebi 2022}
+\end{figure}
 
-\section*{Literature Review}
-Most foreign aid allocated to self-interest and not recipient need (\cite{hoeffler2011})
+%the competition between U.S. and Chinese technology, digital infrastructure spending
+However, the U.S.
+and China do not seem to be cooperating in the effort to improve digital infrastructure worldwide.
+Just months before the PGII announcement, the U.S.
+pioneered a thinly veiled condemnation of China in the Declaration for the Future of the Internet in which more than 60 countries signed a pledge to engage in liberal internet governance (\cite{u.s.departmentofstate2022}).
+Further, policy analysts and scholars have a highlighted a U.S.
+concern over Chinese spending (\cite{hass2021}; \cite{triolo2020}; \cite{hillman2021}).
+But why exactly is the U.S.
+attempting to compete with China in foreign spending and why the digital aspect?
+%need more work here
 
-Aid ineffectiveness in achieving economic growth or promoting democratic institutions (Chong, Gradstein, Calderon)
+\section*{Extant Explanations}
+For a comprehensive but not exhaustive overview of the literature on foreign aid determinants, see Fuchs, Dreher, and Nunnenkamp 2014 \nocite{fuchs2014} (Table 1, p.
+173).
+Simplistically, the foreign aid literature has explained donor behavior as the result of either recipient need or self-interested motivations, with the self-interested reasons being categorizable as political or economic.
+Focusing on the determinants of U.S.
+foreign aid, I argue that the extant explanations do not fully capture either the technology focus of current U.S.
+aid behavior or the antagonistic nature of said behavior.
 
-Aid from U.S. successful in promoting democracy (\cite{blair2022})
+There is evidence that the U.S.
+takes recipient need into account when allocating aid.
+Measured primarily as gross domestic product (GDP) per capita, Hoeffler and Outram (2011) \nocite{hoeffler2011} show that the U.S.
+spends more on poorer countries and that it matters little if there exists human rights abuses or non-democratic rule.
+Theoretically, U.S.
+spending on digital infrastructure could certainly contain some intention to meet recipient need and provide better technology, but it is not clear why the U.S.
+would be be in opposition to China's own digital infrastructure spending if that were the case.
+%mckinlay and buthe
 
-Aid from U.S.
-tends to be more effective in countries that receive no substantial support from China (so we should see funding directed towards countries lacking Chinese aid already)(Dreher et al 2021 AER)
+The traditional account of U.S.
+aid behavior though focuses on self-interested motivations (e.g., \cite{mckinlay1977}).
+Economically, it has been theorized that the U.S.
+gives aid to countries that liberalize trade (\cite{alesina2000}) or export goods to the U.S.
+(\cite{younas2008}; \cite{fuchs2014}).
+It is also often stipulated that the U.S.
+is interested in economic growth (e.g., \cite{bearce2010b}).
+In this lighting, perhaps the U.S.
+is simply keen on stimulating economic growth in recipient countries since there is evidence that the internet is beneficial for developing economies (e.g., see \cite{wallsten2005}).
+However, economic motivations also do not explain why the U.S.
+should be opposed to Chinese spending.
+In fact, it would seem that the U.S.
+should theoretically be in favor of Chinese aid since there is evidence that Chinese aid procures economic growth in recipient countries (\cite{dreher2021}).
 
-Self-interest (Alesina and Dollar)
+Politically, numerous authors highlight how counter-terrorism in the wake of the September 11$^{th}$ attacks resulted in an increase in U.S.
+aid after post-Cold War declines (e.g., \cite{fleck2010}; \cite{bearce2010b}).
+There have been mixed findings on how the democratic status of recipients and how voting records in the United Nations (UN) influences U.S.
+aid.
+Some scholars find a positive relationship between U.S.
+aid and recipient democracy (\cite{alesina2000}) or for the influence of similar UN voting patterns (\cite{alesina2000}; \cite{fuchs2014}), while others find substantially weaker evidence for both (\cite{hoeffler2011}).
+Meanwhile, there is no evidence that the U.S.
+factors in the human rights record of recipients (\cite{hoeffler2011}).
+It is possible that the U.S.
+is returning to a Cold War foreign aid mindset, where the aid they give is entirely concerned with strategically opposing their adversary (\cite{bearce2010b}).
+This would explain the antagonism with China, but not why there should be contention surrounding digital infrastructure and technology.
+%tingley, brech, fleck, dietrich
 
-Chinese aid associated with increase in authoritarian values; World bank aid associated with democracy (Gehring, Kaplan, and Wong)
+\section*{My Explanation}
+I contend that an explanation of current U.S.
+foreign spending behavior that addresses both why it has a strong focus on technology and why this focus seems opposed to China should be based on both political and economic determinants related to technology.
+Specifically, I argue that the U.S.
+has a policy preference for more liberal internet governance, or the rules and norms surrounding the internet in a country,  because a liberal internet enables Western technology to operate within that country.
+That is, I claim that U.S.
+aid can be described by the political determinant of liberal internet governance and the economic determinant of market access for Western technology firms.
 
-aid increases democracy (Dietrich, Wright)
+I will next define internet governance and why it is a political determinant for U.S.
+aid.
+Internet governance was defined in the second phase of the World Summit on the Information Society (WSIS), a meeting led by the UN's International Telecommunication Union (ITU), which was held in Tunis, Tunisia in 2005.
+It is ``the development and application by governments, the private sector and civil society, in their respective roles, of shared principles, norms, rules, decision-making procedures, and programmes that shape the evolution and use of the Internet'' \nocite{worldsummitontheinformationsociety2005} (World Summit on the Information Society 2005, item 34).
+Essentially, internet governance refers to the way in which governments, firms, and citizens interact with the internet.
 
-Commercial self-interest present but possibly overblown (Dreher, Nunnenkamp, Thiele)
+The U.S.
+is concerned about internet governance because there is evidence that the West believes China to be spreading its illiberal model of the internet (e.g., \cite{hillman2021}; \cite{u.s.departmentofstate2022}), or at least that there is a good chance for it to (\cite{triolo2020}).
 
-Bilateral aid more likely to be conditioned on things and less impartial, implying that U.S. should use it more if it is really concerned with internet freedom changes (Christensen, Homer, Nielson)
+However, I argue that the U.S.
+focus on internet governance is not simply about the politics of a liberal internet, such as preferences for democracy or human rights.
+Firstly, according to the literature there is little evidence that U.S.
+government aid discriminates much along the lines of human rights (\cite{hoeffler2011}) or even democracy (\cite{alesina2000}; \cite{hoeffler2011}).
+Rather, I argue that the U.S.'s interest in propping up internet governance around the world chiefly lies in the desire to prevent countries of using the internet in a manner similar to China (\cite{hass2021}), which is more likely to occur with the increase in global Chinese spending.
 
-Trade benefits drive aid (Younas)
+\section*{Implications}
+If U.S.
+foreign spending is truly concerned with internet governance and the economic benefits of inhibiting China's suspected illiberal influence, then it should also be true that the internet governance of potential recipients should be a strong predictor of U.S.
+foreign aid over time.
+This implication leads to the following hypothesis:
+\begin{align*}
+    H_{1}&:\;\text{U.S. aid should be empirically related to recipient internet governance}\\
+\end{align*}
 
-Political economic factors can shape preferences for foreign aid in U.S. (Milner, Tingley)
+\section*{Empirics}
+I test this hypotheses by regressing U.S.
+foreign aid on a measure of internet governance, Chinese aid and controls across 44 African countries between 2010 and 2017.
+The full set of countries includes: Algeria, Angola, Benin, Botswana, Burundi, Cabo Verde, Cameroon, Central African Republic, Chad, Comoros, Republic of the Congo, Cote d'Ivoire, Democratic Republic of the Congo, Djibouti, Egypt, Equatorial Guinea, Ethiopia, Gabon, Gambia, Ghana, Guinea, Guinea-Bissau, Kenya, Lesotho, Liberia, Madagascar, Malawi, Mali, Mauritania, Mauritius, Morocco, Mozambique, Namibia, Niger, Nigeria, Rwanda, Senegal, Sierra Leone, South Africa, Togo, Tunisia, Uganda, Zambia, and Zimbabwe.
 
-\section*{Methodology}
-Dependent variable: US Aid
+I chose to focus on Africa during this time period for three main reasons.
+Firstly, Africa has been a major area of focus for China's finance projects, with the Council on Foreign Relations claiming that ``China already provides more financing for information and communications technology than all multilateral agencies and leading democracies combined do across the continent'' (\cite{kurlantzicketal.2020}).
+Secondly, the data I use for Chinese spending ends in 2017.
+Thirdly, while 2010 is indeed an arbitrary starting point, it seems like a reasonable choice since both Chinese aid and the internet diminish greatly before this period.
 
-Main explanatory variables: Chinese Aid, Internet Freedom
+\subsection*{Data and Variables}
+%internet freedom
+In operationalizing my theory, I measure internet governance as ``internet freedom'' from the Digital Society Project (DSP) index.
+The DSP aims to identify ``how people use social media as a political tool and to explore how political institutions and social media use interact,'' providing data from expert surveys that cover a range of topics such as censorship, disinformation, and monitoring (\cite{mechkova2022}).
+The DSP employs methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}) to try and estimate differences in expert opinion and knowledge.
 
-Controls (according to Cinelli et al.): GDP per capita (recipient need), Internet Usership (Internet development indicators), Political Freedom (possible confounder)
+The DSP is one of two indices measuring a the concept of internet governance and data from the DSP is potentially superior for measuring the concept of internet liberalism in Africa compared to the primary alternative, Freedom on the Net (FOTN, \cite{freedomhouse2022}).
+FOTN frequently features missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.
 
-Chinese data set (Custer et al. 2021; Dreher et al. 2022 (Banking on Beijing))
+%Chinese aid
+The other main explanatory variable is Chinese foreign aid, stemming from the College of William \& Mary's AidData research lab (\cite{custer2021}).
+This is the one of the only datasets on Chinese aid and is certainly the largest and most specific.
+Aid is quantified as official financial and in-kind commitments and is collected using a methodology designed to track under-reported aid flows, such as those from China.
 
-If concerned about simultaneity between US aid and Internet Freedom, maybe use instrument related to Internet Freedom but not to US aid.
+%U.S. aid
+U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{u.s.agencyforinternationaldevelopment2022}). This data was more up-to-date than alternatives.
 
-DiD? With two countries that have similar levels of Internet Freedom and Chinese aid (parallel trends) and then both experience increases in Chinese aid?
-\section*{Complications}
+%controls
+For controls, I include gross domestic product GDP per capita, an indicator for internet development, and a measure for political freedom as controls.
+Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.
+
+GDP per capita is supplied by the United Nations (\cite{unitednationsstatisticsdivision2019}) while the indicator for internet development is provided by the ITU (\cite{internationaltelecommunicationunion2022}).
+This indicator is a proxy for how developed the internet is in a country and represents the percentage of the population using the internet.
+Lastly, the chosen measure for political freedom comes from Polity (\cite{marshall2018}).
+
+Using a GDP deflator (\cite{organizationforeconomicco-operationanddevelopment2022}), the values for U.S. aid, Chinese aid, and GDP per capita are all in constant 2015 USD.
+
+%Missing data
+Of the 10 African countries I am missing data for, the DSP is missing values for Somalia, Eritrea, Sudan, and South Sudan, which might make sense given that Somalia and South Sudan were both at civil war during the observation period while Sudan and Eritrea have experienced instability after previous wars.
+Similarly, Libya was missing data from the ITU following the Western invasion of Libya in 2011.
+However, it is unclear why Sao Tome and Principe and Seychelles are missing Polity data, why Burkina Faso and Eswatani are missing Chinese aid data, or why Tanzania is missing GDP data.
+
+%add discussion of larger descriptive stats table
+Descriptive statistics can be seen in table 2 and 3 below.
+It should be noted that the variable logged Chinese aid is mostly zeroes, as can be seen again in figure 4.
+I performed Grubbs' tests for outliers and indeed, the highest values for logged Chiense aid are outliers from the modal zero.
+All other variables, though, support ample variation and are stationary or trendstationary by themselves.
+
+\subsubsection*{Table 2}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.8]{Figures/628table2.png}
+\end{figure}
+\pagebreak
+
+Description.
+\subsubsection*{Table 3}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.45]{Figures/summary.png}
+\end{figure}
+
+\pagebreak
+A correlation matrix of the variables included in my regressions is given in Figure 2.
+As can be seen, there are a couple relatively strong relationships between the variables.
+The variable for regime type, Polity, has a correlation of 0.62 with the internet freedom variable while logged GDP per capita has a correlation of 0.65 with the variable representing the percent of the population using the internet.
+While possible, I do not think that these correlations are high enough to warrant concerns over multicollinearity or the lack of linear independency in the data.
+Interestingly, logged U.S.
+Aid appears somewhat negatively correlated with logged GDP per capita and variable for the percent using the internet is somewhat positively correlated to that for internet freedom.
+However, all of these correlations are relatively small and should not induce concerns for multicollinearity, although I did not conduct a formal test for multicollinearity.
+
+\pagebreak
+\subsubsection*{Figure 2}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.7]{Figures/628plot3.png}
+\end{figure}
+
+\subsection*{Estimation}
+I chose to use ordinary least squares (OLS) for my estimator while employing two-way fixed effects at the country and year level.
+In choosing fixed effects, I used non-nested model comparison tests that suggested fixed effects to be superior to both no effects and random effects.
+After testing for heteroskedasticity, I included country-clustered standard errors.
+%look at the as.factor(years) as evidence of relationship getting stronger over time?
+
+\pagebreak
+\section*{Results}
+Table 2 displays regression results from a model with and without an interaction between internet freedom and logged Chinese foreign aid.
+\subsection*{Table 4}
+\begin{table}[!htbp] \centering 
+  \caption*{Logged U.S. Aid and Covariates with Two-way Fixed Effects and Clustered SE's on Country}
+  \label{} 
+\begin{tabular}{@{\extracolsep{5pt}}lcc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{2}{c}{Logged U.S. Aid} \\ 
+\cline{2-3} 
+ & Model 1 & Model 2 \\ 
+\hline \\[-1.8ex] 
+ Internet Freedom & 0.021$^{***}$ & 0.021$^{***}$ \\ 
+  & (0.008) & (0.008) \\ 
+  & & \\ 
+ Logged Chinese Aid & 0.006$^{*}$ & 0.006$^{*}$ \\ 
+  & (0.003) & (0.003) \\ 
+  & & \\ 
+ Logged GDP per Capita & 0.045 & 0.045 \\ 
+  & (0.471) & (0.472) \\ 
+  & & \\ 
+ Percent Using Internet & $-$0.013 & $-$0.013 \\ 
+  & (0.009) & (0.009) \\ 
+  & & \\ 
+ Polity & 0.004 & 0.004 \\ 
+  & (0.031) & (0.031) \\ 
+  & & \\ 
+ Internet Freedom $\cdot$ Logged Chinese Aid &  & 0.00000 \\ 
+  &  & (0.0002) \\ 
+  & & \\ 
+\hline \\[-1.8ex] 
+Observations & 352 & 352 \\ 
+R$^{2}$ & 0.038 & 0.038 \\ 
+Adjusted R$^{2}$ & $-$0.140 & $-$0.144 \\ 
+F Statistic & 2.370$^{**}$ (df = 5; 296) & 1.968$^{*}$ (df = 6; 295) \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+\pagebreak
+
+The only covariates that appear statistically significant are internet freedom and Chinese spending.
+Both internet freedom and logged Chinese aid seem to increase logged U.S.
+aid on average when accounting for the other covariates.
+To gauge the substantive size of the coefficients, I plot the marginal effect of internet freedom on logged U.S.
+aid in Figure 3 and the marginal effect of logged Chinese aid on logged U.S.
+aid in Figure 4.
+
+Support for $H_1$?
+
+Interestingly, there is a lack of support for an interaction between internet freedom and logged Chinese aid.
+Apparently, the effects of a recipient's internet freedom level and the amount of Chinese aid they are receiving have completely separate effects on U.S.
+aid.
+However, it is unclear how consequential this lack of support for the interaction term is, since both variables by themselves are significant and two-way fixed effects are used.
+
+\pagebreak
+Figure 3 shows the marginal effect of internet freedom on logged U.S.
+foreign aid.
+The coefficient on the internet freedom variable is 0.0206, implying that a one unit increase in a recipient's internet freedom level is associated with an increase in U.S.
+aid by 2.08 percent on average, accounting for other covariates.
+This appears to be a substantively large effect.
+
+\subsection*{Figure 3}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.7]{Figures/628plot1.png}
+\end{figure}
+
+\pagebreak
+Figure 4 graphs the marginal effect of logged Chinese foreign aid on logged U.S.
+foreign aid.
+The coefficient on the variable for logged Chinese foreign aid is 0.0058, implying that a one percent increase in Chinese aid is associated with a 0.0058 percent increase in U.S.
+aid on average, accounting for other covariates.
+This does not appear to be a substantively large effect.
+
+\subsection*{Figure 4}
+\begin{figure}[htbp]
+    \includegraphics[scale=0.7]{Figures/628plot2.png}
+\end{figure}
+\pagebreak
+
+\section*{Discussion}
+Discuss contribution and implications of support for $H_1$ in light of the literature on foreign aid and LIO.
 
 \section*{Conclusion}
-(Speculate about why US might be rationally interested in increasing internet freedom) Given the literature on the self-interestedness of donors, if U.S.
-aid is changing in response to Chinese aid, what might be a rational motivation?
-Basically: aid effectually increases democracy, democracy is good for business (American Big Tech).
+In this article I argue that . In particular, .
 
+Analyzing 44 African countries from 2010-2017, I find .
+
+I suggest the following future work: Expanding sample to LA Extending years to 2000 Robustness checks with different measurements Greater differntiation between types of aid and money
+
+\nocite{mechkova2022a}
+\nocite{pemstein2022}
+\nocite{coppedge2022}
+\nocite{coppedge2022a}
+\nocite{coppedge2022b}
+\nocite{coppedge2022c}
+\nocite{coppedge2022d}
+\nocite{curini2020}
 \pagebreak

### 2022-12-06 PreviousVersions/OnePager.tex
draft

The People's Republic of China has been an enduring act on the world stage of the last ten years. Unveiling the Belt and Road Initiative (BRI) in 2013 and then the Digital Silk Road (DSR) in 2015, China has become the world's leading financier, surpassing even the United States in yearly foreign aid expenditures (\cite{dreherND}).
Investing in the infrastructure of countries across the globe, it is widely believed that these projects are motivated by both the desire for influence and economic growth, with Chinese companies enjoying unprecedented access to international markets.

Remarkably, the U.S. has decided to act. Surrounded by other members of the G7 at their recent summer summit, President Biden announced that \$600 billion will be spent by the member countries over the next five years to explicitly counter China's intentions (\cite{NPR}). Most interestingly, at least a billion of those dollars are officially earmarked for building internet infrastructure. \nocite{US} \nocite{DSR}

It is not ex ante obvious why the U.S. has taken such a keen and expensive interest in competing with Chinese money or why the internet is a significant focus. China, on the other hand, clearly has much to gain from it's behavior given the domestic importance of economic growth and the lure of becoming a top contender for world power. But does the U.S. really feel a similar demand to grow it's political and financial empire? And even if that is the case, why now?

I argue that the U.S. is trying to combat a perceived increase in digital authoritarianism led by Chinese funding and technology while protecting American Big Tech's access to international markets. I speculate that the U.S. views the use of the internet in other countries as a zero-sum game, where the adoption of Chinese tools and infrastructure prevents U.S.-based companies from flourishing. A clue suggesting this new perspective was a declaration made just two months prior to the G7 spending announcement, where the U.S. and 55 other countries committed to push for a democratic use of the internet worldwide (\cite{Declaration}).

Should this explanation for planned G7 foreign aid be true, it implies that Western money should be spent in an observable pattern consistent with these goals. Specifically, current Chinese foreign aid and the location of profitable technology markets should be strong predictors of future Western foreign aid. The comparison of Western aid over time would thus be the goal of this research, with the expectation that aid increases in countries with strong consumer bases and high exposure to Chinese impact.

The substantive implication of this explanation, should it not be proven false, is the realization that technology and the internet have become central considerations for international politics. Foreign aid, to the extent that it is a strategic endeavor, could be partly determined by cybersecurity concerns and the ability for private firms to profit off of internet activity- motivations that have been largely unconsidered by scholars. A better understanding of these technological incentives might prove salubrious for the predictive power of political science as the internet and AI continue to grow in prominence.

### 2022-12-08 Paper.tex

@@
 \usepackage[utf8]{inputenc}
-\title{\vspace{-2.75cm}Conflating Global Finance with Technological Competition: The U.S.'s Reaction to Chinese Foreign Aid\vspace{-0.5cm}}
+\title{\vspace{-2.75cm}Technological Change and U.S. Foreign Aid\vspace{-0.5cm}}
 \author{Nicholas Ray}
@@
 \begin{abstract}
-    Emphasize policy! \blindtext
+    I argue that the U.S.
+is funding digital infrastructure in Africa to prevent the Chinese from building such infrastructure and closing market access to U.S.
+technology firms.
+Testing the implications that U.S.
+foreign aid spending is directly related to a measure for internet governance and inversely related to Chinese aid, I find mixed results in a sample of 44 African countries from 2010-2017.
+U.S.
+aid seems directly related to both internet governance and Chinese spending.
 \end{abstract}
@@
 \section*{Introduction}
-\footnote{All data and code can be found on GitHub at https://github.com/nnray/Analysis.
-In the ``Code'' folder, navigate to ``628.Rmd'' for my R script.
-My full \LaTeX \;file can be found at https://github.com/nnray/628 under ``Paper.tex''.}
+The United States (U.S.) has been increasingly funding digital infrastructure in countries around the world.\footnote{All R code, data, and \LaTeX \;code can be found at  underlying my analysis can be found on GitHub at https://github.com/nnray.} Simultaneously, the People's Republic of China (China) has advanced its own digital infrastructure projects and financial assistance to become the ``lender of first resort'' for developing countries in the last decade \parencite[1]{dreher2022}.
+Available explanations do not sufficiently address this seeming great power competition along digital infrastructure.
+In this article I put forth an explanation specifying that the U.S.
+is funding digital infrastructure in Africa to prevent the Chinese from building such infrastructure and closing market access to U.S.
+technology firms.
 
-\section*{What Needs Explaining?}
+I make two contributions in the exploration of this topic.
+Firstly, I add to the foreign aid literature by detailing a novel determinant of U.S.
+foreign aid and providing another clue on how Chinese money may have impacted Western donor behavior more generally.
+Secondly, I speak to literature on the liberal international order (LIO) and the concern that China may undermine tenets of the LIO.
+%expand on this a bit?
+
+I proceed as follows.
+In the next section, I introduce puzzling observations of recent U.S.
+behavior followed by a review of current explanations.
+After considering these accounts, I introduce my argument its implications.
+Then, I describe my research design before presenting findings and a discussion of contributions.
+
+\section*{What Needs Explaining}
 %U.S. digital infrastructure spending, other aid related to ICT/internet
@@
 to build digital infrastructure closely resembles the behavior of China over the past two decades.
-Since engaging in their ``going out'' strategy of investing overseas to promote domestic economic growth, China has turned to funding massive infrastructure projects through its Belt and Road Initiative (BRI) and becoming the ``lender of first resort'' for developing countries (\parencite[1]{dreher2022}).
+Since engaging in their ``going out'' strategy of investing overseas to promote domestic economic growth, China has turned to funding massive infrastructure projects through its Belt and Road Initiative (BRI) and becoming the ``lender of first resort'' for developing countries \parencite[1]{dreher2022}.
 Two years after announcing the BRI, China launched the largely symbolic (\cite{tugendhat2021}) Digital Silk Road (DSR) that still encapsulates China's desire to build global digital infrastructure favorable to them (\cite{shen2022}).
@@
 concern over Chinese spending (\cite{hass2021}; \cite{triolo2020}; \cite{hillman2021}).
-But why exactly is the U.S.
-attempting to compete with China in foreign spending and why the digital aspect?
-%need more work here
+Consequently, it appears that the U.S.
+and China are both separately spending on digital infrastructure projects while the U.S.
+actively opposes China's doing so.
 
 \section*{Extant Explanations}
+There are some tentative explanations offered by the global finance literature for this competitive spending behavior surrounding technology.
 For a comprehensive but not exhaustive overview of the literature on foreign aid determinants, see Fuchs, Dreher, and Nunnenkamp 2014 \nocite{fuchs2014} (Table 1, p.
@@
 I contend that an explanation of current U.S.
-foreign spending behavior that addresses both why it has a strong focus on technology and why this focus seems opposed to China should be based on both political and economic determinants related to technology.
-Specifically, I argue that the U.S.
-has a policy preference for more liberal internet governance, or the rules and norms surrounding the internet in a country,  because a liberal internet enables Western technology to operate within that country.
-That is, I claim that U.S.
-aid can be described by the political determinant of liberal internet governance and the economic determinant of market access for Western technology firms.
+foreign spending behavior must address both why it has a strong focus on technology and why this focus seems opposed to China.
+I argue that the U.S.
+is willing to spend aid on providing digital infrastructure because it allows them to build technology in U.S.
+fashion and thereby better protect U.S.
+technology market access to countries liable to Chinese technology influence.
+It is in the U.S.'s interest to attempt technological superiority over the Chinese, which is threatened by China's recent spreading of its technologies and companies to some recipient countries (see \cite{hass2021} for some suggestive anecdotal evidence).
 
-I will next define internet governance and why it is a political determinant for U.S.
-aid.
+If this is true, it should also be true that the U.S.
+is spending more in countries with more liberal internet governance since these are the countries that most allow for the use of U.S.
+technology.
+Further, these countries are likely not being invested in by China, whose aid is suspected on negatively affecting internet governance (e.g., \cite{triolo2020}; \cite{hillman2021}; \cite{u.s.departmentofstate2022}).
+
 Internet governance was defined in the second phase of the World Summit on the Information Society (WSIS), a meeting led by the UN's International Telecommunication Union (ITU), which was held in Tunis, Tunisia in 2005.
@@
 Essentially, internet governance refers to the way in which governments, firms, and citizens interact with the internet.
-
 The U.S.
-is concerned about internet governance because there is evidence that the West believes China to be spreading its illiberal model of the internet (e.g., \cite{hillman2021}; \cite{u.s.departmentofstate2022}), or at least that there is a good chance for it to (\cite{triolo2020}).
-
-However, I argue that the U.S.
-focus on internet governance is not simply about the politics of a liberal internet, such as preferences for democracy or human rights.
-Firstly, according to the literature there is little evidence that U.S.
-government aid discriminates much along the lines of human rights (\cite{hoeffler2011}) or even democracy (\cite{alesina2000}; \cite{hoeffler2011}).
-Rather, I argue that the U.S.'s interest in propping up internet governance around the world chiefly lies in the desire to prevent countries of using the internet in a manner similar to China (\cite{hass2021}), which is more likely to occur with the increase in global Chinese spending.
-
-\section*{Implications}
-If U.S.
-foreign spending is truly concerned with internet governance and the economic benefits of inhibiting China's suspected illiberal influence, then it should also be true that the internet governance of potential recipients should be a strong predictor of U.S.
-foreign aid over time.
-This implication leads to the following hypothesis:
-\begin{align*}
-    H_{1}&:\;\text{U.S. aid should be empirically related to recipient internet governance}\\
-\end{align*}
+should be expected to allocate more aid to countries with higher measures of internet governance if they are truly bolstering U.S.
+firm access to technology markets.
 
 \section*{Empirics}
-I test this hypotheses by regressing U.S.
+Restating the implications of my argument, U.S.
+foreign aid should be directly related to recipient country internet governance since the U.S.
+is trying to build infrastructure that allows for an economically and politically liberal environment for U.S.
+technology to thrive.
+These same environments are most likely to exist in a country not financed by China, since China is sharing technology along with its money that could negatively affect recipient internet governance.
+Thus I will evaluate the following hypotheses:
+
+\begin{align*}
+    H_{1}&:\;\text{U.S. aid should be directly related to recipient internet governance}\\
+    H_{2}&:\;\text{U.S. aid should be inversely related to Chinese aid}\\
+\end{align*}
+I test these hypotheses by regressing U.S.
 foreign aid on a measure of internet governance, Chinese aid and controls across 44 African countries between 2010 and 2017.
@@
 \subsection*{Data and Variables}
-%internet freedom
-In operationalizing my theory, I measure internet governance as ``internet freedom'' from the Digital Society Project (DSP) index.
+%internet governance
+In operationalizing my theory, I measure internet governance using the Digital Society Project (DSP) index.
 The DSP aims to identify ``how people use social media as a political tool and to explore how political institutions and social media use interact,'' providing data from expert surveys that cover a range of topics such as censorship, disinformation, and monitoring (\cite{mechkova2022}).
@@
 For controls, I include gross domestic product GDP per capita, an indicator for internet development, and a measure for political freedom as controls.
-Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.
+Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet governance.
 
@@
 
-Description.
 \subsubsection*{Table 3}
@@
 As can be seen, there are a couple relatively strong relationships between the variables.
-The variable for regime type, Polity, has a correlation of 0.62 with the internet freedom variable while logged GDP per capita has a correlation of 0.65 with the variable representing the percent of the population using the internet.
+The variable for regime type, Polity, has a correlation of 0.62 with the internet governance variable while logged GDP per capita has a correlation of 0.65 with the variable representing the percent of the population using the internet.
 While possible, I do not think that these correlations are high enough to warrant concerns over multicollinearity or the lack of linear independency in the data.
 Interestingly, logged U.S.
-Aid appears somewhat negatively correlated with logged GDP per capita and variable for the percent using the internet is somewhat positively correlated to that for internet freedom.
+Aid appears somewhat negatively correlated with logged GDP per capita and variable for the percent using the internet is somewhat positively correlated to that for internet governance.
 However, all of these correlations are relatively small and should not induce concerns for multicollinearity, although I did not conduct a formal test for multicollinearity.
@@
 \section*{Results}
-Table 2 displays regression results from a model with and without an interaction between internet freedom and logged Chinese foreign aid.
-\subsection*{Table 4}
+Table 2 displays regression results from a model with and without an interaction between internet governance and logged Chinese foreign aid.
 \begin{table}[!htbp] \centering 
-  \caption*{Logged U.S. Aid and Covariates with Two-way Fixed Effects and Clustered SE's on Country}
+  \caption{Logged U.S. Aid and Covariates with Two-way Fixed Effects and Clustered SE's on Country} 
   \label{} 
@@
 \hline \\[-1.8ex] 
- Internet Freedom & 0.021$^{***}$ & 0.021$^{***}$ \\ 
+ Internet Governance & 0.021$^{***}$ & 0.021$^{***}$ \\ 
   & (0.008) & (0.008) \\ 
@@
   & & \\ 
- Internet Freedom $\cdot$ Logged Chinese Aid &  & 0.00000 \\ 
+ Internet Governance * Logged Chinese Aid &  & 0.00000 \\ 
   &  & (0.0002) \\ 
@@
 
-The only covariates that appear statistically significant are internet freedom and Chinese spending.
-Both internet freedom and logged Chinese aid seem to increase logged U.S.
+The only covariates that appear statistically significant are internet governance and Chinese spending.
+Both internet governance and logged Chinese aid seem to increase logged U.S.
 aid on average when accounting for the other covariates.
-To gauge the substantive size of the coefficients, I plot the marginal effect of internet freedom on logged U.S.
+To gauge the substantive size of the coefficients, I plot the marginal effect of internet governance on logged U.S.
 aid in Figure 3 and the marginal effect of logged Chinese aid on logged U.S.
-aid in Figure 4.
+aid in Figure 4, both from model 1 (seen in Table 2).
 
-Support for $H_1$?
+According to Table 2 and Figures 3 and 4, there is support for $H_1$.
+$H_2$ is contradicted by the evidence, meanwhile.
+However, I would be cautious about inferences from the results pertaining to logged Chinese aid as most of the values are zero, as can be seen in Figure 4.
+It is thus likely that the downward sloping line is overestimated due to all the zeros.
 
-Interestingly, there is a lack of support for an interaction between internet freedom and logged Chinese aid.
-Apparently, the effects of a recipient's internet freedom level and the amount of Chinese aid they are receiving have completely separate effects on U.S.
-aid.
-However, it is unclear how consequential this lack of support for the interaction term is, since both variables by themselves are significant and two-way fixed effects are used.
+I also tested for an interaction between internet governance and logged Chinese aid, which was not part of my theory. The interaction term is not significant according to standard levels (p=0.05).
 
 \pagebreak
-Figure 3 shows the marginal effect of internet freedom on logged U.S.
+Figure 3 shows the marginal effect of internet governance on logged U.S.
 foreign aid.
-The coefficient on the internet freedom variable is 0.0206, implying that a one unit increase in a recipient's internet freedom level is associated with an increase in U.S.
+The coefficient on the internet governance variable is 0.0206, implying that a one unit increase in a recipient's internet governance level is associated with an increase in U.S.
 aid by 2.08 percent on average, accounting for other covariates.
@@
 \section*{Discussion}
-Discuss contribution and implications of support for $H_1$ in light of the literature on foreign aid and LIO.
+The support of $H_1$ but not $H_2$ indicate mixed results for my theory.
+However, it seems that robustness checks with mulitple other measures would be required to make more inferences about U.S.
+behavior, including for regime type and especially Chinese aid.
+The variable for Chinese aid contains mostly zeros with a few, large values that are qualified as outliers.
+
+To the extent that my theory is supported, I contribute to the foreign aid literature by detailing a potentially new determinant for U.S.
+behavior.
+The internet and technology as a foreign policy and geopolitical issue will only continue to rise, and it's important that scholarly theories keep pace with contemporary observation.
+Secondly, some international relations literature has expressed concerns that China will undermine common tenets of the LIO, such as economic or political liberalism (e.g., \cite{weiss2021a}).
+While I do not study China's behavior here, my analysis indicate somewhat that the U.S.
+behaves as if China is indeed damaging political and economic liberalism, especially as it relates to the liberal use of the internet.
 
 \section*{Conclusion}
-In this article I argue that . In particular, .
+In this article I argue that the U.S.
+is funding digital infrastructure in Africa to prevent the Chinese from building such infrastructure and closing market access to U.S.
+technology firms.
+I test the implications that U.S.
+foreign aid spending should be directly related to a measure for internet governance, or the rules and norms surrounding the internet, and inversely related to Chinese aid spending.
 
-Analyzing 44 African countries from 2010-2017, I find .
+Analyzing 44 African countries from 2010-2017, I find mixed results for the theory, as U.S. aid indeed seems directly related to internet governance but also directly related to Chinese spending.
 
-I suggest the following future work: Expanding sample to LA Extending years to 2000 Robustness checks with different measurements Greater differntiation between types of aid and money
+In future work, other measures for Chinese spending and influence would be critical to analyzing their digital infrastructure activities in Africa that go poorly reflected in aid data.
+The deriving of other implications and more falsification attempts are also necessary to test this explanation for recent U.S.
+spending behavior.
 

### 2022-12-13 Paper.tex
hyperref

@@
 \doublespacing
+\usepackage[colorlinks,citecolor=cyan,urlcolor=blue,bookmarks=false,hypertexnames=true]{hyperref}
 \usepackage{footmisc}
@@
 }
+\DeclareCiteCommand{\citeyear}
+    {\usebibmacro{prenote}}
+    {\bibhyperref{\printfield{year}}\bibhyperref{\printfield{extrayear}}}
+    {\multicitedelim}
+    {\usebibmacro{postnote}}
+
+\DeclareCiteCommand{\citeyearpar}[\mkbibparens]
+    {\usebibmacro{prenote}}
+    {\bibhyperref{\printfield{year}}\bibhyperref{\printfield{extrayear}}}
+    {\multicitedelim}
+    {\usebibmacro{postnote}}
 \addbibresource{ForeignAidBib.bib}
@@
 Two years after announcing the BRI, China launched the largely symbolic (\cite{tugendhat2021}) Digital Silk Road (DSR) that still encapsulates China's desire to build global digital infrastructure favorable to them (\cite{shen2022}).
-Some examples of relevant Chinese projects are shown in table 1 \nocite{agbebi2022}.
+Some examples of relevant Chinese projects are shown in table 1 (from \cite{agbebi2022}).
 
@@
 There are some tentative explanations offered by the global finance literature for this competitive spending behavior surrounding technology.
-For a comprehensive but not exhaustive overview of the literature on foreign aid determinants, see Fuchs, Dreher, and Nunnenkamp 2014 \nocite{fuchs2014} (Table 1, p.
+For a comprehensive but not exhaustive overview of the literature on foreign aid determinants, see Fuchs, Dreher, and Nunnenkamp (\citeyear{fuchs2014} p.
 173).
@@
 takes recipient need into account when allocating aid.
-Measured primarily as gross domestic product (GDP) per capita, Hoeffler and Outram (2011) \nocite{hoeffler2011} show that the U.S.
+Measured primarily as gross domestic product (GDP) per capita, Hoeffler and Outram (\citeyear{hoeffler2011}) show that the U.S.
 spends more on poorer countries and that it matters little if there exists human rights abuses or non-democratic rule.
@@
 Internet governance was defined in the second phase of the World Summit on the Information Society (WSIS), a meeting led by the UN's International Telecommunication Union (ITU), which was held in Tunis, Tunisia in 2005.
-It is ``the development and application by governments, the private sector and civil society, in their respective roles, of shared principles, norms, rules, decision-making procedures, and programmes that shape the evolution and use of the Internet'' \nocite{worldsummitontheinformationsociety2005} (World Summit on the Information Society 2005, item 34).
+It is ``the development and application by governments, the private sector and civil society, in their respective roles, of shared principles, norms, rules, decision-making procedures, and programmes that shape the evolution and use of the Internet'' (\cite{worldsummitontheinformationsociety2005}, item 34).
 Essentially, internet governance refers to the way in which governments, firms, and citizens interact with the internet.

### 2022-12-14 Paper.tex
update

@@
 \section*{Introduction}
-The United States (U.S.) has been increasingly funding digital infrastructure in countries around the world.\footnote{All R code, data, and \LaTeX \;code can be found at  underlying my analysis can be found on GitHub at https://github.com/nnray.} Simultaneously, the People's Republic of China (China) has advanced its own digital infrastructure projects and financial assistance to become the ``lender of first resort'' for developing countries in the last decade \parencite[1]{dreher2022}.
+The United States (U.S.) has been increasingly funding digital infrastructure in countries around the world.\footnote{All R code, data, and \LaTeX \;code underlying my analysis can be found in the corresponding repository on GitHub at https://github.com/nnray.} Simultaneously, the People's Republic of China (China) has advanced its own digital infrastructure projects and financial assistance to become the ``lender of first resort'' for developing countries in the last decade \parencite[1]{dreher2022}.
 Available explanations do not sufficiently address this seeming great power competition along digital infrastructure.

### 2022-12-20 Other/PreviousVersions/FivePager.tex
Preamble

\section*{Introduction}
The United States (U.S.) has been openly concerned about the internet since at least 1997, when President Bill Clinton exhorted the world to adopt a laissez-faire, commercially friendly approach to the internet (\cite{government1997}). By the 2000's, these preferences for a ``free'' internet were officially crystallized into foreign policy initiatives as so-called ``internet freedom'' programs cropped up in different facets of the U.S. government, such as the Department of State (e.g., \cite{henry2014}; \cite{government2021a}) and the U.S. Agency for Global Media (\cite{government2022}).

U.S. attention to internet freedom has continued to grow (e.g., \cite{government2010}), crescendoing most recently in 2022. All in one year, bipartisan legislation was introduced in the Senate to increase funding to internet freedom programs (\cite{government2022a}), a declaration was signed by the U.S. and over 60 other countries to combat global ``digital authoritarianism'' (\cite{government2022b}), and an unprecedented, U.S.-led G7 spending plan was unveiled to improve physical and digital infrastructure abroad (\cite{government2022c}).

Along the way, the People's Republic of China (China) and the Russian Federation (Russia) have been named the antagonists to global internet freedom and the explicit motivations behind the U.S.'s focus on the subject (see \cite{government2010}, \cite{government2022b}). During this intensifying period of U.S. apprehension of apparent digital authoritarianism, China launched it's Belt and Road Initiative (BRI) in 2013 and the accompanying Digital Silk Road Initiative (DSR) in 2015. Collectively, these projects have catalyzed over one trillion U.S. dollars in developing-country investment with the intent to improve China's perception and influence in the world while fostering it's domestic economic growth (\cite{dreher2022}).

Ample academic work has recently attempted to outline the effects of China's aid program (e.g., \cite{blair2021}; \cite{eichenauer2021}; \cite{dreher2018}). Yet, little research has explored the potential effects of China's spending on U.S. foreign policy. In this paper, I seek to understand if U.S. foreign aid, an important aspect of foreign policy, can be understood as a response to China's aid ambitions.

Literature shows that foreign aid is typically a self-interested tool (e.g., \cite{hoeffler2011}) and that U.S. aid can be effective in stimulating liberal democratic values in a contemporary environment saturated with Chinese aid (\cite{blair2022}). Given these findings, it seems likely that a view of U.S. foreign aid over time would reflect the historical policy importance of internet freedom and the U.S.'s desire to oppose both China's growing influence and their authoritarian model of the internet. Specifically, I hypothesize that U.S. foreign aid to Africa is related to a recipient's internet freedom level and their influx of Chinese aid.

Should this relationship be borne out in the data, future work would have to precisely explain how the U.S. benefits from higher global levels of internet freedom if foreign aid is indeed a self-interested tool. If no relationship is apparent, however, it must similarly be explained why the U.S. is failing to reconcile their foreign policy position on internet freedom with their intention to compete with China's expanding aid project since there is evidence that foreign aid can be effective in promoting liberal democratic values.

Understanding the U.S.'s actualization of internet preferences has important implications not only for the foreign aid literature, but also for international relations and comparative scholars more broadly. Firstly, any result that indicated the importance of internet considerations in inter-state behavior would be the first of its kind to my knowledge, and would signify a potentially novel dimension along which states interact. Secondly, this paper would help inform the dynamics of the U.S. and China's ongoing contest for economic and political influence by highlighting another issue area by which they compete. Lastly, this work would raise questions about the generalizability of state concern over the internet, which could encompass studies on conflict and cyberspace, research on democratization and information asymmetries, and political economic work on globalization.

I proceed by highlighting the relevant literature on foreign aid before moving on to the operationalization and testing of the hypothesis. In the concluding remarks, I offer tentative theories for various anticipated results.

\section*{Literature Review}
\subsection*{Altruistic versus Self-Interested Aid}
Scholars have long debated if donors of foreign aid are motivated by altruism or self-interest. The received wisdom on foreign aid states that donors often give aid according to their strategic concerns (e.g., \cite{mckinlay1977}; \cite{alesina2000}), and economic trade has more recently been shown to be a predictor of donor aid. Specifically, donors seem to distribute aid amongst recipients that import goods that donors have a comparative advantage in producing (\cite{younas2008}).

However, some studies have implied altruism is also an important factor in determining aid flows, especially for private aid and non-governmental organizations (\cite{buthe2012}). Other scholars have analyzed the potential for aid to reduce poverty, concluding that poverty reduction is occurring but in an inefficient manner (\cite{collier2002}). The consensus, though, still seems to be that recipient humanitarian need is largely dominated by donor self-interest (e.g., \cite{hoeffler2011}).

The balance between strategic interests and altruism has additionally been conceived of along the lines of donor maturity, with more established donors potentially distributing aid in more or less altruistic ways than ``new'' donors (\cite{dreher2011}). Generally, though, it is found that traditional aid providers care less for recipient need than previously thought while relatively young donors are less self-interested than expected by previous literature.

\subsection*{Foreign Aid: An Effective Tool?}
Regardless of the exact source of motivation behind foreign aid, there has been a significant amount of disagreement over whether foreign aid is effective in achieving any objectives. In some cases, for example, it appears that the lack of discernible and consistent motivations is the very reason why aid can be found to be ineffective (\cite{collier2002}). Aid has also been found to be ineffective in altering levels of inequality or poverty in recipient countries, even if those countries have relatively strong institutions (\cite{chong2009a}).

There is, however, modest evidence that economic aid effectively contributes to democratic consolidation in Africa (\cite{dietrich2015a}). Effectiveness can vary by certain characteristics of aid as well, with bilateral aid being more likely to be conditioned on a recipient country's level of corruption. Multilateral aid, on the other hand, is generally more impartial and unconditional, resulting in diminished effectiveness when compared to bilateral aid (\cite{christensen2011}).

Interestingly, evidence for the effectiveness of aid appears to have strengthened with China's monumental entrance into the aid arena. For example, when comparing the effects of Chinese and U.S. aid in Africa, U.S. aid is found to be relatively successful in promoting liberal democratic values amongst Afrobarometer respondents (\cite{blair2022}). Chinese aid, by contrast, largely has null or even similarly positive effects on democratic values. In a separate study relating World Bank and Chinese aid, foreign aid from the World Bank is associated with the same liberal democratic values and even improved country stability. A feeling of acceptance towards autocratic leadership is found to be connected to Chinese aid (\cite{gehring2022}).

Yet there is a substantial amount of nuance to aid effectiveness. Aid from China is definitively linked with positive economic growth in recipient countries, for example, while U.S. aid is most effective in countries not currently receiving large amounts of Chinese aid (\cite{dreher2021}).

Thus, it can be gathered from previous literature that donor behavior, including that of the U.S., should be expected to be a product of strategic considerations. Furthermore, while not always effective in securing these considerations, it is plausible that the current atmosphere of aid-giving in Africa represents a scenario in which aid tied to liberal democratic ideals can be effectual.

\section*{Empirics}
\subsection*{Data and Variables}
In approaching an analysis of the relationship between foreign aid and internet freedom, it perhaps makes most sense to firstly define the main explanatory variable of internet freedom. The Digital Society Project (DSP) aims to identify ``how people use social media as a political tool and to explore how political institutions and social media use interact,'' providing data from expert surveys that cover a range of topics such as censorship, disinformation, and monitoring (\cite{mechkova2022}). The DSP employs methodology designed by the Varieties of Democracy project (V-Dem, \cite{coppedge2022}) to try and estimate differences in expert opinion and knowledge.

Data from the DSP is potentially superior for measuring the concept of internet freedom compared to the primary alternative, Freedom on the Net (FOTN, \cite{house2022}). FOTN features frequent missing data for the continent of Africa and does not take into account possible variations in expert opinion and knowledge like the DSP does.

The other main explanatory variable is Chinese foreign aid, stemming from the College of William \& Mary's AidData research lab (\cite{custer2021}). This is the one of the only datasets on Chinese aid and is certainly the largest and most specific. Aid is quantified as official financial and in-kind commitments and is measured in constant 2017 USD.

U.S. foreign aid is the dependent variable, measured as fiscal-year disbursements (\cite{government2022e}). It is measured in constant 2022 USD.

The starting year for the holistic dataset is 2000, defined by the lower bound of data availability for internet freedom, and 2017 is the ending year per data availability for Chinese foreign aid. There is some evidence of left-truncation or missing data in the information on internet freedom, but otherwise all African countries are accounted for.

\subsection*{Methodology}
To estimate the association between U.S. foreign aid and the main explanatory variables, I plan to use ordinary least squares (OLS) with country fixed-effects. Fixed-effects at the country level help account for any underlying similarity between countries that could induce spuriousness.

In addition to including fixed-effects, there are multiple controls that seem appropriate to incorporate. According to the most recent advice for determining good controls (\cite{cinelli2022}), I include gross domestic product (GDP) per capita, an indicator for internet development, and political freedom as controls. Each of these phenomena are possible confounders, being theoretically related to at least one of the types of aid and internet freedom.

GDP per capita is assessed in constant 2021 USD and as is supplied by the World Bank (\cite{bank2022}). The indicator for internet development, internet usage by country, is calculated as a percentage and is provided by the International Telecommunication Union (\cite{itu2022}). Lastly, the chosen measure for political freedom comes from V-Dem (\cite{coppedge2022a}).

\section*{Limitations}
There are at least a couple important limitations to be clear about. Firstly, there is undoubtedly some measurement error in the quantification of internet freedom brought about due to possible differences in the idea of internet freedom employed by the U.S. and DSP. In other words, even if the U.S.'s conception of internet freedom was constant over time, the DSP definition is likely to be somewhat distinct.

Secondly, there is potential simultaneity between U.S. foreign aid and internet freedom. More advanced econometric modeling techniques, such as difference-in-differences (DiD) or instrumental variables regression (IV) may be imperative here.

\section*{Conclusion}
It is possible that there is no meaningful association between U.S. foreign aid and internet freedom. The lack of association would in itself present a sort of puzzle, though, since the U.S. clearly has a foreign policy interest in maintaining internet freedom and a seemingly effective tool of foreign aid to pursue that interest. It may take time, however, for the ever-increasing advances in U.S. financial commitment to internet freedom to be reflected in spending data.

If there is indeed a relationship, I speculate that the U.S. is interested in keeping the internet free because it is good for business. There is evidence that China is sharing it's technology, like censorship and monitoring tools, that could allow more authoritarian-leaning developing countries to prevent the use of U.S. technology and to strictly allow for Chinese substitutes instead. These developing countries would then harness the economic benefits of the internet, like job growth, while constraining potentially negative political externalities such as the free assembly of anti-incumbent online groups. China would stand to gain domestic economic growth as it's technology firms expand beyond it's borders, and everybody wins- except the U.S.

### 2022-12-20 Other/PreviousVersions/Lightning.tex
Preamble

Scholars have long debated over the determinants of foreign aid and its effectiveness, generally concluding that foreign aid is a strategic tool for donors (e.g., \cite{alesina2000}; \cite{hoeffler2011}) and that its effectiveness greatly varies (e.g., see \cite{christensen2011}). Most recently, there has been a resurgence in the exploration of aid effectiveness due to an unprecedented availability of Chinese foreign aid. With regard to the U.S., for example, researchers have found that U.S. aid has become more effective at stimulating liberal democratic values (\cite{blair2022}), especially in countries not receiving large amounts of Chinese aid (\cite{dreher2021}).

However, there has been a lack of equal reconsideration for how Chinese aid potentially influences the strategic aims of donors. I argue that the abundance of Chinese money should alter donor strategy, especially for the U.S. Specifically, if the U.S. is a rational actor, it should place a greater emphasis on a recipient's ``internet freedom'' level when allocating aid in light of China's rising spending influence. This is rational because 1) advancing global internet freedom in opposition to China's model of the internet has been a vocal aspect of U.S. foreign policy for decades (e.g., \cite{government2010}, \cite{government2022b}) and 2) promoting internet freedom could be an effective use of resources given the findings on U.S. aid's ability to encourage liberal values in the contemporary aid environment (\cite{blair2022}).

This logic leads to the following hypothesis:
\begin{align*}
    H1 &:\text{U.S. aid should be related to a recipient's internet freedom level and the amount of Chinese}\\
    &\text{aid they receive.}
\end{align*}

I plan to use OLS with country-year fixed effects to evaluate how much of the variation in U.S. foreign aid to Africa from 2010 to 2017 can be explained by an interaction between Chinese foreign aid and recipient internet freedom levels. Internet freedom is measured by the Digital Society Project (DSP), using V-Dem methodology and expert-coded surveys to analyze how social media is used as a political tool by countries around the world (\cite{mechkova2022}). Important controls include a country's GDP per capita and an indicator for internet development.
